---
title: "Compartment Tracking"
output: html_document
date: "2026-08-04"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(ggrepel)
library(readxl)
library(tidyverse)
library(writexl)

```

# Import Data

## Blood intervention effects

The blood workbook contains the adjusted, standardized average treatment effects used in `RESULTS_for_Trenton_blood_milk_analysis.docx`. FDR is recalculated within each compartment and timepoint, as described in that document.

```{r}

blood_file <-
  "../1. Data/blood_mummichog_input_for_trenton.xlsx"

blood_sheets <-
  blood_file %>%
  excel_sheets()

blood_results <-
  blood_sheets %>%
  set_names() %>%
  map_dfr(
    ~ blood_file %>%
      read_excel(
        sheet = .x,
        .name_repair = "unique_quiet"
      ) %>%
      mutate(modality = .x)
  )

```

## Milk intervention effects

```{r}

milk_results_raw <-
  readRDS(
    "../1. Data/combined_intervention_effects_results_combined_arms.RDS"
  )

milk_alignment <-
  read_csv(
    paste0(
      "/Users/trentondailey-chwalibog/My Drive/9 ugent/",
      "imicPaper/results/IMiC_alignment.csv"
    ),
    show_col_types = FALSE,
    name_repair = "unique_quiet"
  )

```

## BEP supplement data

```{r}

supplement_results_raw <-
  read_csv(
    paste0(
      "/Users/trentondailey-chwalibog/My Drive/9 ugent/imicPaper/",
      "data/additional datasets/ProcessedDataMISAME3_VAMS.csv"
    ),
    show_col_types = FALSE,
    name_repair = "unique_quiet"
  )

supplement_columns <-
  supplement_results_raw %>%
  names() %>%
  str_subset("^supplement;APSE")

length(supplement_columns)

```

## Inspect imported data

```{r}

blood_results

glimpse(blood_results)

blood_results %>%
  count(
    modality,
    compartment,
    timepoint,
    name = "number_of_features"
  ) %>%
  arrange(
    modality,
    timepoint
  ) %>%
  print(n = Inf)

```

# Wrangle Data

## Calculate FDR and classify effects

```{r}

fdr_threshold <- 0.05
mass_tolerance_ppm <- 25

blood_results <-
  blood_results %>%
  mutate(
    fdr =
      p.adjust(
        raw_p,
        method = "BH"
      ),
    .by =
      c(
        modality,
        timepoint,
        contrast
      )
  ) %>%
  mutate(
    compartment =
      recode(
        modality,
        "MaternalPlasma" = "Maternal plasma",
        "VamsPostnatalMaternal" = "Maternal VAMS",
        "VamsPostnatalInfant" = "Infant VAMS"
      ),
    direction =
      case_when(
        effect_size < 0 ~ "Downregulated",
        effect_size > 0 ~ "Upregulated",
        TRUE ~ "No change"
      ),
    significant = fdr < fdr_threshold,
    significance =
      case_when(
        significant ~ "FDR significant",
        TRUE ~ "Not FDR significant"
      ),
    logp =
      -log10(
        pmax(
          raw_p,
          .Machine$double.xmin
        )
      ),
    timepoint =
      factor(
        timepoint,
        levels =
          c(
            "incl",
            "tri3",
            "acco",
            "pn12",
            "pn34",
            "pn56"
          )
      )
  ) %>%
  arrange(
    compartment,
    timepoint,
    raw_p,
    feature
  )

```

## Prepare milk results

Only MISAME-III untargeted milk features are retained because the blood and APSE supplement data are from MISAME-III. The supplied milk `sigFDR` classification is preserved.

```{r}

milk_alignment <-
  milk_alignment %>%
  transmute(
    feature_key =
      str_to_lower(mtb_id_MISAME3),
    mz = mz_MISAME3,
    rt = rt_MISAME3
  ) %>%
  distinct(
    feature_key,
    .keep_all = TRUE
  )

milk_results <-
  milk_results_raw %>%
  filter(
    study == "Misame",
    contrast == "BEP",
    measure == "ATE",
    str_detect(
      biomarker,
      regex(
        "^rlc_",
        ignore_case = TRUE
      )
    )
  ) %>%
  mutate(
    feature_key =
      str_to_lower(biomarker)
  ) %>%
  left_join(
    milk_alignment,
    by = "feature_key"
  ) %>%
  transmute(
    compartment = "Milk",
    modality = "Milk",
    timepoint =
      recode(
        visit,
        "14-21 days" = "1421d",
        "1-2 mo." = "pn12",
        "3-4 mo." = "pn34"
      ),
    contrast,
    feature = label_f,
    mz,
    rt,
    ion_mode =
      case_when(
        str_detect(feature_key, "_pos_") ~ "positive",
        str_detect(feature_key, "_neg_") ~ "negative",
        TRUE ~ NA_character_
      ),
    effect_size = est,
    direction =
      case_when(
        est < 0 ~ "Downregulated",
        est > 0 ~ "Upregulated",
        TRUE ~ "No change"
      ),
    raw_p = pval,
    fdr = pval_adj,
    significance =
      case_when(
        sigFDR == 1 ~ "FDR significant",
        TRUE ~ "Not FDR significant"
      ),
    significant = sigFDR == 1,
    logp =
      -log10(
        pmax(
          raw_p,
          .Machine$double.xmin
        )
      )
  )

```

## Combine compartments

```{r}

compartment_results <-
  bind_rows(
    blood_results,
    milk_results
  ) %>%
  mutate(
    compartment =
      factor(
        compartment,
        levels =
          c(
            "Maternal plasma",
            "Maternal VAMS",
            "Milk",
            "Infant VAMS"
          )
      )
  ) %>%
  arrange(
    compartment,
    timepoint,
    raw_p,
    feature
  )

compartment_results

glimpse(compartment_results)

```

## Prepare APSE supplement results

The APSE data contain 18 supplement replicates. A feature is reliably detected when present in at least half of those replicates. Supplement-abundant features have mean abundance at least one standard deviation above the mean across supplement features.

```{r}

supplement_results <-
  supplement_results_raw %>%
  select(
    MZ,
    RT,
    Metabolite_Feature_Label,
    all_of(supplement_columns)
  ) %>%
  pivot_longer(
    cols = all_of(supplement_columns),
    names_to = "supplement_sample",
    values_to = "supplement_abundance"
  ) %>%
  group_by(
    MZ,
    RT,
    Metabolite_Feature_Label
  ) %>%
  summarise(
    supplement_replicates = n(),
    detected_replicates =
      sum(
        !is.na(supplement_abundance) &
        supplement_abundance > 0
      ),
    mean_supplement_abundance =
      mean(
        supplement_abundance,
        na.rm = TRUE
      ),
    median_supplement_abundance =
      median(
        supplement_abundance,
        na.rm = TRUE
      ),
    sd_supplement_abundance =
      sd(
        supplement_abundance,
        na.rm = TRUE
      ),
    .groups = "drop"
  ) %>%
  mutate(
    across(
      c(
        mean_supplement_abundance,
        median_supplement_abundance,
        sd_supplement_abundance
      ),
      ~ replace(
        .x,
        is.nan(.x),
        NA_real_
      )
    ),
    supplement_feature = Metabolite_Feature_Label,
    supplement_feature_key =
      str_to_lower(supplement_feature),
    supplement_mz = MZ,
    supplement_rt = RT,
    ion_mode =
      case_when(
        str_detect(supplement_feature_key, "_pos_") ~ "positive",
        str_detect(supplement_feature_key, "_neg_") ~ "negative",
        TRUE ~ NA_character_
      ),
    supplement_detected =
      detected_replicates > 0,
    supplement_reliably_detected =
      detected_replicates >=
      ceiling(supplement_replicates / 2)
  )

supplement_abundance_threshold <-
  supplement_results %>%
  summarise(
    threshold =
      mean(
        mean_supplement_abundance,
        na.rm = TRUE
      ) +
      sd(
        mean_supplement_abundance,
        na.rm = TRUE
      )
  ) %>%
  pull(threshold)

supplement_results <-
  supplement_results %>%
  mutate(
    supplement_abundant =
      supplement_reliably_detected &
      mean_supplement_abundance >=
      supplement_abundance_threshold
  ) %>%
  select(
    supplement_feature,
    supplement_feature_key,
    supplement_mz,
    supplement_rt,
    ion_mode,
    supplement_replicates,
    detected_replicates,
    mean_supplement_abundance,
    median_supplement_abundance,
    sd_supplement_abundance,
    supplement_detected,
    supplement_reliably_detected,
    supplement_abundant
  )

supplement_results

```

## Isolate statistically significant effects

```{r}

significant_compartment_results <-
  compartment_results %>%
  filter(
    significant,
    direction %in%
      c(
        "Downregulated",
        "Upregulated"
      )
  )

significant_compartment_results

significant_compartment_results %>%
  count(
    compartment,
    timepoint,
    direction,
    name = "number_of_features"
  ) %>%
  arrange(
    compartment,
    timepoint,
    direction
  ) %>%
  print(n = Inf)

```

## Validate results against the supplied document

```{r}

reported_significant_counts <-
  tribble(
    ~compartment, ~timepoint, ~reported_number,
    "Maternal plasma", "incl", 17,
    "Maternal plasma", "tri3", 10,
    "Maternal plasma", "pn12", 593,
    "Maternal VAMS", "tri3", 0,
    "Maternal VAMS", "pn56", 18,
    "Infant VAMS", "acco", 0,
    "Infant VAMS", "pn12", 3,
    "Infant VAMS", "pn34", 8,
    "Infant VAMS", "pn56", 18
  ) %>%
  mutate(
    timepoint =
      factor(
        timepoint,
        levels = levels(blood_results$timepoint)
      )
  )

significant_count_validation <-
  blood_results %>%
  summarise(
    calculated_number =
      sum(
        fdr < fdr_threshold,
        na.rm = TRUE
      ),
    .by =
      c(
        compartment,
        timepoint
      )
  ) %>%
  right_join(
    reported_significant_counts,
    by =
      c(
        "compartment",
        "timepoint"
      )
  ) %>%
  mutate(
    calculated_number =
      replace_na(
        calculated_number,
        0L
      ),
    matches_document =
      calculated_number == reported_number
  ) %>%
  arrange(
    compartment,
    timepoint
  )

significant_count_validation

stopifnot(
  all(significant_count_validation$matches_document)
)

```

## Validate milk and supplement data

```{r}

milk_count_validation <-
  milk_results %>%
  count(
    timepoint,
    direction,
    significant,
    name = "number_of_features"
  ) %>%
  arrange(
    timepoint,
    desc(significant),
    direction
  )

milk_count_validation %>%
  print(n = Inf)

milk_alignment_validation <-
  milk_results %>%
  summarise(
    number_of_features = n_distinct(feature),
    features_missing_mz =
      n_distinct(
        feature[is.na(mz)]
      ),
    features_missing_rt =
      n_distinct(
        feature[is.na(rt)]
      )
  )

milk_alignment_validation

stopifnot(
  n_distinct(milk_results$feature) == 37070,
  length(supplement_columns) == 18
)

```

# Analysis

## Compare statistically significant and non-significant features

```{r}

significance_comparison <-
  compartment_results %>%
  filter(
    direction %in%
      c(
        "Downregulated",
        "Upregulated"
      )
  ) %>%
  count(
    compartment,
    timepoint,
    direction,
    significance,
    name = "number_of_features"
  ) %>%
  group_by(
    compartment,
    timepoint,
    direction
  ) %>%
  mutate(
    total_features =
      sum(number_of_features),
    percent_of_features =
      100 * number_of_features / total_features
  ) %>%
  ungroup() %>%
  arrange(
    compartment,
    timepoint,
    direction,
    desc(significance)
  )

significance_comparison %>%
  print(n = Inf)

```

## Export statistically significant effects

```{r}

compartment_results_root <-
  "../3. Results/Compartment Tracking"

dir.create(
  compartment_results_root,
  recursive = TRUE,
  showWarnings = FALSE
)

significant_compartment_results %>%
  group_by(
    compartment,
    timepoint,
    direction
  ) %>%
  group_walk(
    ~ .x %>%
      write_xlsx(
        path =
          file.path(
            compartment_results_root,
            paste0(
              str_to_lower(.y$compartment) %>%
                str_replace_all("[^a-z0-9]+", "_"),
              "_",
              .y$timepoint,
              "_",
              str_to_lower(.y$direction),
              ".xlsx"
            )
          )
      ),
    .keep = TRUE
  )

write_xlsx(
  list(
    significant_features = significant_compartment_results,
    significance_comparison = significance_comparison,
    validation = significant_count_validation
  ),
  path =
    file.path(
      compartment_results_root,
      "compartment_tracking_significant_features.xlsx"
    )
)

```

# Original Volcano Plots

## Prepare volcano-plot data

```{r}

volcano_colors <-
  c(
    "FDR significant" = "#F2AE2E",
    "Raw P < 0.05" = "#56B4E9",
    "Not significant" = "black"
  )

volcano_plot_data <-
  compartment_results %>%
  mutate(
    volcano_group =
      case_when(
        significant ~ "FDR significant",
        raw_p < 0.05 ~ "Raw P < 0.05",
        TRUE ~ "Not significant"
      ),
    original_label =
      case_when(
        significant ~ feature,
        TRUE ~ NA_character_
      )
  )

```

## Maternal plasma

```{r}

maternal_plasma_volcano_plot <-
  volcano_plot_data %>%
  filter(compartment == "Maternal plasma") %>%
  ggplot(
    aes(
      x = effect_size,
      y = logp,
      color = volcano_group
    )
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.4
  ) +
  geom_point(
    alpha = 0.65,
    size = 1
  ) +
  geom_text_repel(
    aes(label = original_label),
    color = "grey45",
    size = 2.5,
    max.overlaps = 10,
    na.rm = TRUE
  ) +
  facet_grid(
    rows = vars(timepoint),
    scales = "free_y"
  ) +
  scale_color_manual(values = volcano_colors) +
  labs(
    title = "Maternal plasma (combined arms, adjusted)",
    x = "Average Treatment Effect",
    y = expression(-log[10]("raw P-value")),
    color = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "bottom")

maternal_plasma_volcano_plot

```

## Postnatal maternal VAMS

```{r}

maternal_vams_volcano_plot <-
  volcano_plot_data %>%
  filter(compartment == "Maternal VAMS") %>%
  ggplot(
    aes(
      x = effect_size,
      y = logp,
      color = volcano_group
    )
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.4
  ) +
  geom_point(
    alpha = 0.65,
    size = 1
  ) +
  geom_text_repel(
    aes(label = original_label),
    color = "grey45",
    size = 2.5,
    max.overlaps = 18,
    na.rm = TRUE
  ) +
  facet_grid(
    rows = vars(timepoint),
    scales = "free_y"
  ) +
  scale_color_manual(values = volcano_colors) +
  labs(
    title = "Postnatal maternal VAMS (combined arms, adjusted)",
    x = "Average Treatment Effect",
    y = expression(-log[10]("raw P-value")),
    color = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "bottom")

maternal_vams_volcano_plot

```

## Postnatal infant VAMS

```{r}

infant_vams_volcano_plot <-
  volcano_plot_data %>%
  filter(compartment == "Infant VAMS") %>%
  ggplot(
    aes(
      x = effect_size,
      y = logp,
      color = volcano_group
    )
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.4
  ) +
  geom_point(
    alpha = 0.65,
    size = 1
  ) +
  geom_text_repel(
    aes(label = original_label),
    color = "grey45",
    size = 2.5,
    max.overlaps = 18,
    na.rm = TRUE
  ) +
  facet_grid(
    rows = vars(timepoint),
    scales = "free_y"
  ) +
  scale_color_manual(values = volcano_colors) +
  labs(
    title = "Postnatal infant VAMS (combined arms, adjusted)",
    x = "Average Treatment Effect",
    y = expression(-log[10]("raw P-value")),
    color = NULL
  ) +
  theme_bw(base_size = 8) +
  theme(legend.position = "bottom")

infant_vams_volcano_plot

```

# Mummichog

Mummichog is run separately for each compartment, timepoint, ionization mode, and direction. All measured features remain in the background. For a directional run, features pointing in the other direction are assigned a P-value of 1. The command follows the supplied analysis document: 10-ppm pathway matching, `human_mfn`, and a raw P-value cutoff of 0.05.

## Set Mummichog paths

```{r}

mummichog_exec <-
  "/Users/trentondailey-chwalibog/anaconda3/envs/biospe/bin/mummichog"

mummichog_root <-
  file.path(
    compartment_results_root,
    "Mummichog"
  )

dir.create(
  mummichog_root,
  recursive = TRUE,
  showWarnings = FALSE
)

stopifnot(
  file.exists(mummichog_exec)
)

```

## Prepare Mummichog jobs

```{r}

mummichog_jobs <-
  compartment_results %>%
  filter(
    !is.na(mz),
    !is.na(rt),
    !is.na(raw_p),
    ion_mode %in%
      c(
        "positive",
        "negative"
      )
  ) %>%
  distinct(
    compartment,
    timepoint,
    ion_mode
  ) %>%
  crossing(
    analysis =
      c(
        "overall",
        "upregulated",
        "downregulated"
      )
  ) %>%
  mutate(
    job_id =
      paste(
        compartment,
        timepoint,
        ion_mode,
        analysis,
        sep = "_"
      ) %>%
      str_to_lower() %>%
      str_replace_all("[^a-z0-9]+", "_")
  ) %>%
  arrange(
    compartment,
    timepoint,
    ion_mode,
    analysis
  )

mummichog_jobs

```

## Run Mummichog

```{r}

mummichog_jobs %>%
  pwalk(
    function(
      compartment,
      timepoint,
      ion_mode,
      analysis,
      job_id
    ) {
      mummichog_input <-
        compartment_results %>%
        filter(
          .data$compartment == .env$compartment,
          .data$timepoint == .env$timepoint,
          .data$ion_mode == .env$ion_mode,
          !is.na(mz),
          !is.na(rt),
          !is.na(raw_p)
        ) %>%
        arrange(
          raw_p,
          feature
        ) %>%
        mutate(
          input_row = row_number(),
          mummichog_p =
            case_when(
              analysis == "upregulated" & effect_size < 0 ~ 1,
              analysis == "downregulated" & effect_size > 0 ~ 1,
              TRUE ~ pmax(
                raw_p,
                .Machine$double.xmin
              )
            ),
          statistic =
            -log10(mummichog_p)
        )

      mummichog_input %>%
        select(
          mz,
          rt,
          mummichog_p,
          statistic
        ) %>%
        write_tsv(
          file.path(
            mummichog_root,
            paste0(job_id, "_input.txt")
          ),
          col_names = FALSE
        )

      mummichog_input %>%
        select(
          input_row,
          compartment,
          timepoint,
          ion_mode,
          feature,
          mz,
          rt,
          effect_size,
          raw_p,
          fdr,
          direction
        ) %>%
        write_csv(
          file.path(
            mummichog_root,
            paste0(job_id, "_input_key.csv")
          )
        )

      existing_output <-
        mummichog_root %>%
        list.dirs(
          recursive = FALSE,
          full.names = TRUE
        ) %>%
        str_subset(
          fixed(paste0(".", job_id))
        )

      if (length(existing_output) == 0) {
        withr::with_dir(
          mummichog_root,
          system2(
            command = mummichog_exec,
            args =
              c(
                "-f",
                paste0(job_id, "_input.txt"),
                "-o",
                job_id,
                "-m",
                ion_mode,
                "-u",
                "10",
                "-n",
                "human_mfn",
                "-c",
                "0.05"
              )
          )
        )
      }
    }
  )

```

## Import Mummichog pathway results

```{r}

mummichog_output_folders <-
  mummichog_jobs %>%
  mutate(
    output_folder =
      map_chr(
        job_id,
        ~ mummichog_root %>%
          list.dirs(
            recursive = FALSE,
            full.names = TRUE
          ) %>%
          str_subset(
            fixed(paste0(".", .x))
          ) %>%
          sort() %>%
          last()
      )
  )

mummichog_pathway_results <-
  mummichog_output_folders %>%
  mutate(
    data =
      map2(
        output_folder,
        job_id,
        ~ file.path(
          .x,
          "tables",
          paste0(
            "mcg_pathwayanalysis_",
            .y,
            ".tsv"
          )
        ) %>%
          read_tsv(
            show_col_types = FALSE,
            name_repair = "unique_quiet"
          )
      )
  ) %>%
  select(
    compartment,
    timepoint,
    ion_mode,
    analysis,
    job_id,
    data
  ) %>%
  unnest(cols = data) %>%
  rename(raw_p = `p-value`) %>%
  mutate(
    fdr =
      p.adjust(
        raw_p,
        method = "BH"
      ),
    .by = job_id
  ) %>%
  arrange(
    compartment,
    timepoint,
    ion_mode,
    analysis,
    raw_p,
    pathway
  )

mummichog_pathway_results

```

# Putative Feature Names

Mummichog names are putative mass-to-compound candidates. A single feature can have several candidates, so the candidate count is retained and the first candidate is used only as a concise plotting label.

## Import Mummichog candidate names

```{r}

## Import Mummichog candidate names

mummichog_name_map <-
  mummichog_output_folders %>%
  filter(analysis == "overall") %>%
  mutate(
    candidate_file =
      file.path(
        output_folder,
        "tables",
        "userInput_to_EmpiricalCompounds.tsv"
      ),
    input_key_file =
      file.path(
        mummichog_root,
        paste0(
          job_id,
          "_input_key.csv"
        )
      ),
    candidates =
      map(
        candidate_file,
        ~ read_tsv(
          .x,
          show_col_types = FALSE,
          name_repair = "unique_quiet"
        ) %>%
          transmute(
            input_row =
              `input_row...1` %>%
              str_remove("^row") %>%
              as.integer(),
            compound_ids = compounds,
            putative_names = compound_names
          ) %>%
          filter(
            !is.na(putative_names),
            putative_names != ""
          ) %>%
          group_by(input_row) %>%
          summarise(
            compound_ids =
              compound_ids %>%
              unique() %>%
              paste(collapse = ";"),
            putative_names =
              putative_names %>%
              unique() %>%
              paste(collapse = "$"),
            .groups = "drop"
          )
      ),
    input_key =
      map(
        input_key_file,
        ~ read_csv(
          .x,
          show_col_types = FALSE,
          col_types =
            cols(
              .default = col_guess(),
              compartment = col_character(),
              timepoint = col_character(),
              ion_mode = col_character(),
              feature = col_character(),
              direction = col_character()
            )
        )
      ),
    data =
      map2(
        candidates,
        input_key,
        ~ .y %>%
          left_join(
            .x,
            by = "input_row"
          )
      )
  ) %>%
  select(data) %>%
  unnest(cols = data) %>%
  filter(!is.na(putative_names)) %>%
  mutate(
    putative_name =
      putative_names %>%
      str_split_fixed(
        "[$;]",
        2
      ) %>%
      .[, 1] %>%
      str_squish(),
    candidate_count =
      putative_names %>%
      str_split("[$;]") %>%
      map_int(
        ~ .x %>%
          str_squish() %>%
          discard(~ .x == "") %>%
          n_distinct()
      )
  ) %>%
  distinct(
    compartment,
    timepoint,
    ion_mode,
    feature,
    .keep_all = TRUE
  )

mummichog_name_map

```

## Add putative names to all results

```{r}

named_compartment_results <-
  compartment_results %>%
  left_join(
    mummichog_name_map %>%
      select(
        compartment,
        timepoint,
        ion_mode,
        feature,
        compound_ids,
        putative_names,
        putative_name,
        candidate_count
      ),
    by =
      c(
        "compartment",
        "timepoint",
        "ion_mode",
        "feature"
      )
  ) %>%
  mutate(
    feature_name =
      coalesce(
        putative_name,
        feature
      ),
    normalized_name =
      putative_name %>%
      str_to_lower() %>%
      str_replace_all("[^a-z0-9]+", ""),
    feature_label =
      case_when(
        significant & !is.na(putative_name) ~
          paste0(
            putative_name,
            " (",
            feature,
            ")"
          ),
        significant ~ feature,
        TRUE ~ NA_character_
      )
  )

named_compartment_results %>%
  filter(significant) %>%
  arrange(
    compartment,
    timepoint,
    raw_p
  )

```

# Volcano Plots With Putative Names

```{r}

named_volcano_plots <-
  named_compartment_results %>%
  mutate(
    volcano_group =
      case_when(
        significant ~ "FDR significant",
        raw_p < 0.05 ~ "Raw P < 0.05",
        TRUE ~ "Not significant"
      )
  ) %>%
  group_split(compartment) %>%
  set_names(
    named_compartment_results %>%
      distinct(compartment) %>%
      pull(compartment)
  ) %>%
  map(
    ~ .x %>%
      ggplot(
        aes(
          x = effect_size,
          y = logp,
          color = volcano_group
        )
      ) +
      geom_vline(
        xintercept = 0,
        linetype = "dashed",
        linewidth = 0.4
      ) +
      geom_point(
        alpha = 0.65,
        size = 1
      ) +
      geom_text_repel(
        aes(label = feature_label),
        color = "grey35",
        size = 2.3,
        max.overlaps = 20,
        na.rm = TRUE
      ) +
      facet_grid(
        rows = vars(timepoint),
        scales = "free_y"
      ) +
      scale_color_manual(values = volcano_colors) +
      labs(
        title =
          paste0(
            unique(.x$compartment),
            " (combined arms, adjusted)"
          ),
        x = "Average Treatment Effect",
        y = expression(-log[10]("raw P-value")),
        color = NULL
      ) +
      theme_bw(base_size = 8) +
      theme(legend.position = "bottom")
  )

named_volcano_plots

```

# Identify Features in the BEP Supplement

Significant compartment features are matched to APSE supplement features by exact identifier where available, followed by same-ion-mode mass matching within 25 ppm. The closest reliably detected candidate is retained, while the number of possible candidates is reported.

## Match significant features to the supplement

```{r}

significant_named_results <-
  named_compartment_results %>%
  filter(significant) %>%
  mutate(
    compartment = as.character(compartment),
    timepoint = as.character(timepoint),
    tracking_id =
      paste(
        compartment,
        timepoint,
        feature,
        sep = "__"
      ),
    feature_key =
      str_to_lower(feature),
    mz_lower =
      mz *
      (1 - mass_tolerance_ppm / 1000000),
    mz_upper =
      mz *
      (1 + mass_tolerance_ppm / 1000000)
  )

supplement_matches_all <-
  significant_named_results %>%
  left_join(
    supplement_results,
    by =
      join_by(
        ion_mode,
        mz_lower <= supplement_mz,
        mz_upper >= supplement_mz
      ),
    relationship = "many-to-many"
  ) %>%
  mutate(
    supplement_ppm_difference =
      abs(mz - supplement_mz) /
      ((mz + supplement_mz) / 2) *
      1000000,
    supplement_rt_difference =
      abs(rt - supplement_rt),
    exact_supplement_feature_match =
      feature_key == supplement_feature_key,
    supplement_candidate_count =
      sum(
        !is.na(supplement_feature),
        na.rm = TRUE
      ),
    .by = tracking_id
  )

supplement_matches <-
  supplement_matches_all %>%
  arrange(
    tracking_id,
    desc(exact_supplement_feature_match),
    desc(supplement_reliably_detected),
    supplement_ppm_difference,
    supplement_rt_difference
  ) %>%
  distinct(
    tracking_id,
    .keep_all = TRUE
  ) %>%
  mutate(
    supplement_match_type =
      case_when(
        is.na(supplement_feature) ~ "No match within 25 ppm",
        exact_supplement_feature_match ~ "Exact feature identifier",
        TRUE ~ "Mass within 25 ppm"
      ),
    supplement_status =
      case_when(
        is.na(supplement_feature) ~ "No matched supplement feature",
        supplement_abundant ~ "Supplement-abundant",
        supplement_reliably_detected ~ "Reliably detected",
        supplement_detected ~ "Detected",
        TRUE ~ "Not detected"
      )
  )

supplement_matches %>%
  count(
    compartment,
    timepoint,
    direction,
    supplement_status,
    name = "number_of_features"
  ) %>%
  arrange(
    compartment,
    timepoint,
    direction,
    supplement_status
  ) %>%
  print(n = Inf)

```

## Add supplement results to named features

```{r}

named_compartment_results <-
  named_compartment_results %>%
  mutate(
    tracking_id =
      paste(
        compartment,
        timepoint,
        feature,
        sep = "__"
      )
  ) %>%
  left_join(
    supplement_matches %>%
      select(
        tracking_id,
        supplement_feature,
        supplement_mz,
        supplement_rt,
        supplement_ppm_difference,
        supplement_rt_difference,
        supplement_candidate_count,
        supplement_match_type,
        supplement_replicates,
        detected_replicates,
        mean_supplement_abundance,
        median_supplement_abundance,
        supplement_detected,
        supplement_reliably_detected,
        supplement_abundant,
        supplement_status
      ),
    by = "tracking_id"
  )

```

# Track Features Across Compartments

Cross-compartment comparisons use direction only because absolute intensities are not comparable across platforms. Maternal and infant VAMS are first linked by exact feature identifier. Remaining features are linked by an identical normalized putative name or by ionization mode and mass within 25 ppm. Name-based matches remain putative and mass matches can include chance matches.

## Identify matching features

```{r}

significant_named_results <-
  named_compartment_results %>%
  filter(significant) %>%
  mutate(
    compartment = as.character(compartment),
    timepoint = as.character(timepoint),
    tracking_id =
      paste(
        compartment,
        timepoint,
        feature,
        sep = "__"
      )
  )

compartment_pairs <-
  significant_named_results %>%
  select(
    tracking_id,
    compartment,
    timepoint,
    ion_mode,
    feature,
    mz,
    rt,
    effect_size,
    direction,
    fdr,
    putative_name,
    normalized_name
  ) %>%
  rename_with(
    ~ paste0(.x, "_1")
  ) %>%
  crossing(
    significant_named_results %>%
      select(
        tracking_id,
        compartment,
        timepoint,
        ion_mode,
        feature,
        mz,
        rt,
        effect_size,
        direction,
        fdr,
        putative_name,
        normalized_name
      ) %>%
      rename_with(
        ~ paste0(.x, "_2")
      )
  ) %>%
  filter(
    compartment_1 < compartment_2
  ) %>%
  mutate(
    ppm_difference =
      abs(mz_1 - mz_2) /
      ((mz_1 + mz_2) / 2) *
      1000000,
    exact_feature_match =
      feature_1 == feature_2,
    mass_match =
      ion_mode_1 == ion_mode_2 &
      ppm_difference <= 25,
    similar_name_match =
      !is.na(normalized_name_1) &
      !is.na(normalized_name_2) &
      normalized_name_1 == normalized_name_2,
    same_direction =
      direction_1 == direction_2,
    match_type =
      case_when(
        exact_feature_match ~ "Exact feature identifier",
        similar_name_match ~ "Same putative name",
        mass_match ~ "Mass within 25 ppm",
        TRUE ~ NA_character_
      )
  ) %>%
  filter(!is.na(match_type)) %>%
  arrange(
    desc(exact_feature_match),
    ppm_difference,
    fdr_1,
    fdr_2
  ) %>%
  distinct(
    tracking_id_1,
    tracking_id_2,
    .keep_all = TRUE
  )

compartment_pairs

```

## Summarise concordance

```{r}

compartment_concordance <-
  compartment_pairs %>%
  count(
    compartment_1,
    timepoint_1,
    compartment_2,
    timepoint_2,
    match_type,
    same_direction,
    name = "number_of_matches"
  ) %>%
  group_by(
    compartment_1,
    timepoint_1,
    compartment_2,
    timepoint_2,
    match_type
  ) %>%
  mutate(
    percent_of_matches =
      100 * number_of_matches / sum(number_of_matches)
  ) %>%
  ungroup() %>%
  arrange(
    compartment_1,
    timepoint_1,
    compartment_2,
    timepoint_2,
    match_type,
    desc(same_direction)
  )

compartment_concordance %>%
  print(n = Inf)

```

## Prepare compartment-tracking figure

```{r}

tracked_feature_ids <-
  compartment_pairs %>%
  filter(same_direction) %>%
  select(
    tracking_id_1,
    tracking_id_2
  ) %>%
  pivot_longer(
    cols = everything(),
    values_to = "tracking_id"
  ) %>%
  distinct(tracking_id)

compartment_tracking_plot_data <-
  significant_named_results %>%
  semi_join(
    tracked_feature_ids,
    by = "tracking_id"
  ) %>%
  mutate(
    compartment_timepoint =
      paste(
        compartment,
        timepoint,
        sep = " · "
      ),
    compartment_timepoint =
      factor(
        compartment_timepoint,
        levels =
          c(
            "Maternal plasma · incl",
            "Maternal plasma · tri3",
            "Maternal plasma · pn12",
            "Maternal VAMS · tri3",
            "Maternal VAMS · pn56",
            "Milk · 1421d",
            "Milk · pn12",
            "Milk · pn34",
            "Infant VAMS · acco",
            "Infant VAMS · pn12",
            "Infant VAMS · pn34",
            "Infant VAMS · pn56"
          ),
        labels =
          c(
            "Maternal plasma · Enrollment",
            "Maternal plasma · Third trimester",
            "Maternal plasma · 1–2 months",
            "Maternal VAMS · Third trimester",
            "Maternal VAMS · 5–6 months",
            "Milk · 14–21 days",
            "Milk · 1–2 months",
            "Milk · 3–4 months",
            "Infant VAMS · Birth",
            "Infant VAMS · 1–2 months",
            "Infant VAMS · 3–4 months",
            "Infant VAMS · 5–6 months"
          )
      ),
    tracking_name =
      coalesce(
        putative_name,
        paste0(
          "m/z ",
          round(mz, 4)
        )
      ),
    supplement_label =
      case_when(
        supplement_abundant ~ "Supplement-abundant",
        supplement_reliably_detected ~ "Detected in supplement",
        TRUE ~ "Not reliably detected in supplement"
      )
  )

```

## Generate compartment-tracking figure

```{r}

compartment_tracking_plot <-
  compartment_tracking_plot_data %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name,
      color = direction
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey65",
    alpha = 0.6,
    linewidth = 0.5
  ) +
  geom_point(
    size = 2,
    alpha = 0.9
  ) +
  geom_text_repel(
    aes(label = tracking_name),
    size = 2.3,
    color = "grey30",
    max.overlaps = 30,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values =
      c(
        "Downregulated" = "#4E79A7",
        "Upregulated" = "#E15759"
      )
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    color = "Direction"
  ) +
  theme_bw(base_size = 8) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

compartment_tracking_plot

```

```{r}
# -------------------------------------------------------------------
# Upregulated features
# -------------------------------------------------------------------

compartment_tracking_upregulated_plot <-
  compartment_tracking_plot_data %>%
  filter(direction == "Upregulated") %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey65",
    alpha = 0.6,
    linewidth = 0.5
  ) +
  geom_point(
    aes(shape = supplement_label),
    color = "#E15759",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_text_repel(
    aes(label = tracking_name),
    size = 2.3,
    color = "grey30",
    box.padding = 0.5,
    point.padding = 0.3,
    segment.color = "grey60",
    segment.size = 0.3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  scale_shape_manual(
    values =
      c(
        "Supplement-abundant" = 17,
        "Detected in supplement" = 15,
        "Not reliably detected in supplement" = 16
      )
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    title = "Upregulated Features",
    shape = "BEP Supplement"
  ) +
  theme_bw(base_size = 8) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

compartment_tracking_upregulated_plot

# -------------------------------------------------------------------
# Downregulated features
# -------------------------------------------------------------------

compartment_tracking_downregulated_plot <-
  compartment_tracking_plot_data %>%
  filter(direction == "Downregulated") %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey65",
    alpha = 0.6,
    linewidth = 0.5
  ) +
  geom_point(
    aes(shape = supplement_label),
    color = "#4E79A7",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_text_repel(
    aes(label = tracking_name),
    size = 2.3,
    color = "grey30",
    box.padding = 0.5,
    point.padding = 0.3,
    segment.color = "grey60",
    segment.size = 0.3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  scale_shape_manual(
    values =
      c(
        "Supplement-abundant" = 17,
        "Detected in supplement" = 15,
        "Not reliably detected in supplement" = 16
      )
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    title = "Downregulated Features",
    shape = "BEP Supplement"
  ) +
  theme_bw(base_size = 8) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

compartment_tracking_downregulated_plot

# -------------------------------------------------------------------
# Export figures
# -------------------------------------------------------------------

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "compartment_tracking_upregulated_plot.png"
    ),
  plot = compartment_tracking_upregulated_plot,
  width = 14,
  height = 14,
  units = "in",
  dpi = 600,
  bg = "white"
)

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "compartment_tracking_downregulated_plot.png"
    ),
  plot = compartment_tracking_downregulated_plot,
  width = 14,
  height = 14,
  units = "in",
  dpi = 600,
  bg = "white"
)
```
# Export Results

```{r}

write_xlsx(
  list(
    named_features = named_compartment_results,
    significant_named_features =
      named_compartment_results %>%
      filter(significant),
    significance_comparison = significance_comparison,
    supplement_feature_summary = supplement_results,
    supplement_matches = supplement_matches,
    supplement_candidates = supplement_matches_all,
    matched_features = compartment_pairs,
    concordance = compartment_concordance,
    mummichog_pathways = mummichog_pathway_results,
    blood_validation = significant_count_validation,
    milk_validation = milk_count_validation
  ),
  path =
    file.path(
      compartment_results_root,
      "compartment_tracking_final_results.xlsx"
    )
)

walk2(
  named_volcano_plots,
  names(named_volcano_plots),
  ~ ggsave(
    filename =
      file.path(
        compartment_results_root,
        paste0(
          str_to_lower(.y) %>%
            str_replace_all("[^a-z0-9]+", "_"),
          "_named_volcano_plot.png"
        )
      ),
    plot = .x,
    width = 9,
    height = 12,
    units = "in",
    dpi = 600,
    bg = "white"
  )
)

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "compartment_tracking_plot.png"
    ),
  plot = compartment_tracking_plot,
  width = 12,
  height = 6.75,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

# Linked Features Only

## Upregulated Linked Features

```{r}

linked_upregulated_ids <-
  compartment_pairs %>%
  filter(
    same_direction,
    direction_1 == "Upregulated",
    direction_2 == "Upregulated"
  ) %>%
  select(
    tracking_id_1,
    tracking_id_2
  ) %>%
  pivot_longer(
    cols = everything(),
    values_to = "tracking_id"
  ) %>%
  distinct(tracking_id)

linked_upregulated_features <-
  named_compartment_results %>%
  semi_join(
    linked_upregulated_ids,
    by = "tracking_id"
  ) %>%
  filter(direction == "Upregulated") %>%
  arrange(
    compartment,
    timepoint,
    putative_name,
    feature
  )

linked_upregulated_features

linked_upregulated_features %>%
  count(
    compartment,
    timepoint,
    name = "linked_features"
  ) %>%
  arrange(
    compartment,
    timepoint
  ) %>%
  print(n = Inf)

```
## Downregulated Linked Features

```{r}

linked_downregulated_ids <-
  compartment_pairs %>%
  filter(
    same_direction,
    direction_1 == "Downregulated",
    direction_2 == "Downregulated"
  ) %>%
  select(
    tracking_id_1,
    tracking_id_2
  ) %>%
  pivot_longer(
    cols = everything(),
    values_to = "tracking_id"
  ) %>%
  distinct(tracking_id)

linked_downregulated_features <-
  named_compartment_results %>%
  semi_join(
    linked_downregulated_ids,
    by = "tracking_id"
  ) %>%
  filter(direction == "Downregulated") %>%
  arrange(
    compartment,
    timepoint,
    putative_name,
    feature
  )

linked_downregulated_features

linked_downregulated_features %>%
  count(
    compartment,
    timepoint,
    name = "linked_features"
  ) %>%
  arrange(
    compartment,
    timepoint
  ) %>%
  print(n = Inf)

```

## Plot Linked Features

### Upregulated

```{r}

linked_upregulated_plot <-
  compartment_tracking_plot_data %>%
  semi_join(
    linked_upregulated_ids,
    by = "tracking_id"
  ) %>%
  filter(direction == "Upregulated") %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey65",
    alpha = 0.6,
    linewidth = 0.5
  ) +
  geom_point(
    aes(shape = supplement_label),
    color = "#E15759",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_text_repel(
    aes(label = tracking_name),
    size = 2.3,
    color = "grey30",
    box.padding = 0.5,
    point.padding = 0.3,
    segment.color = "grey60",
    segment.size = 0.3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  scale_shape_manual(
    values =
      c(
        "Supplement-abundant" = 17,
        "Detected in supplement" = 15,
        "Not reliably detected in supplement" = 16
      )
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    title = "Linked Upregulated Features",
    shape = "BEP Supplement"
  ) +
  theme_bw(base_size = 8) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

linked_upregulated_plot

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "linked_upregulated_features.png"
    ),
  plot = linked_upregulated_plot,
  width = 14,
  height = 14,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

### Downregulated

```{r}

linked_downregulated_plot <-
  compartment_tracking_plot_data %>%
  semi_join(
    linked_downregulated_ids,
    by = "tracking_id"
  ) %>%
  filter(direction == "Downregulated") %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey65",
    alpha = 0.6,
    linewidth = 0.5
  ) +
  geom_point(
    aes(shape = supplement_label),
    color = "#4E79A7",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_text_repel(
    aes(label = tracking_name),
    size = 2.3,
    color = "grey30",
    box.padding = 0.5,
    point.padding = 0.3,
    segment.color = "grey60",
    segment.size = 0.3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  scale_shape_manual(
    values =
      c(
        "Supplement-abundant" = 17,
        "Detected in supplement" = 15,
        "Not reliably detected in supplement" = 16
      )
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    title = "Linked Downregulated Features",
    shape = "BEP Supplement"
  ) +
  theme_bw(base_size = 8) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

linked_downregulated_plot

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "linked_downregulated_features.png"
    ),
  plot = linked_downregulated_plot,
  width = 14,
  height = 14,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

# In Supplement Bar Plot

```{r}

# -------------------------------------------------------------------
# Prepare supplement-status plot data
# -------------------------------------------------------------------

supplement_status_plot_data <-
  supplement_matches %>%
  count(
    compartment,
    timepoint,
    direction,
    supplement_status,
    name = "number_of_features"
  ) %>%
  group_by(
    compartment,
    timepoint,
    direction
  ) %>%
  mutate(
    total_features =
      sum(number_of_features),
    percent_features =
      100 *
      number_of_features /
      total_features,
    segment_label =
      case_when(
        percent_features >= 4 ~
          paste0(
            number_of_features,
            "\n",
            round(
              percent_features,
              1
            ),
            "%"
          ),
        TRUE ~ ""
      )
  ) %>%
  ungroup() %>%
  mutate(
    compartment_timepoint =
      paste(
        compartment,
        timepoint,
        sep = " · "
      ),
    compartment_timepoint =
      factor(
        compartment_timepoint,
        levels =
          c(
            "Maternal plasma · incl",
            "Maternal plasma · tri3",
            "Maternal plasma · pn12",
            "Maternal VAMS · pn56",
            "Milk · 1421d",
            "Milk · pn12",
            "Milk · pn34",
            "Infant VAMS · pn12",
            "Infant VAMS · pn34",
            "Infant VAMS · pn56"
          ),
        labels =
          c(
            "Maternal plasma · Enrollment",
            "Maternal plasma · Third trimester",
            "Maternal plasma · 1–2 months",
            "Maternal VAMS · 5–6 months",
            "Milk · 14–21 days",
            "Milk · 1–2 months",
            "Milk · 3–4 months",
            "Infant VAMS · 1–2 months",
            "Infant VAMS · 3–4 months",
            "Infant VAMS · 5–6 months"
          )
      ),
    supplement_status =
      factor(
        supplement_status,
        levels =
          c(
            "No matched supplement feature",
            "Not detected",
            "Detected",
            "Reliably detected",
            "Supplement-abundant"
          )
      ),
    direction =
      factor(
        direction,
        levels =
          c(
            "Upregulated",
            "Downregulated"
          )
      )
  )

# -------------------------------------------------------------------
# Colours
# -------------------------------------------------------------------

supplement_status_colors <-
  c(
    "No matched supplement feature" = "#D9D9D9",
    "Not detected" = "#969696",
    "Detected" = "#F2CC8F",
    "Reliably detected" = "#59A89C",
    "Supplement-abundant" = "#7B2CBF"
  )

# -------------------------------------------------------------------
# Plot
# -------------------------------------------------------------------

supplement_status_plot <-
  supplement_status_plot_data %>%
  ggplot(
    aes(
      x = number_of_features,
      y = compartment_timepoint,
      fill = supplement_status
    )
  ) +
  geom_col(
    position = "fill",
    width = 0.75,
    color = "white",
    linewidth = 0.25
  ) +
  geom_text(
    aes(label = segment_label),
    position =
      position_fill(
        vjust = 0.5
      ),
    size = 2.3,
    color = "black"
  ) +
  facet_wrap(
    vars(direction),
    ncol = 1
  ) +
  scale_x_continuous(
    labels = scales::percent,
    expand =
      expansion(
        mult =
          c(
            0,
            0.01
          )
      )
  ) +
  scale_fill_manual(
    values = supplement_status_colors,
    drop = FALSE
  ) +
  labs(
    x = "Percentage of FDR-Significant Features",
    y = NULL,
    fill = "BEP Supplement Status"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(
      size = 8
    ),
    legend.position = "bottom",
    legend.box = "vertical"
  ) +
  guides(
    fill =
      guide_legend(
        nrow = 2,
        byrow = TRUE
      )
  )

supplement_status_plot

# -------------------------------------------------------------------
# Export
# -------------------------------------------------------------------

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "supplement_status_by_compartment.png"
    ),
  plot = supplement_status_plot,
  width = 12,
  height = 10,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

# Prepare cross-compartment plotting data

```{r}

# -------------------------------------------------------------------
# Retain metabolites matched across at least two positions
# -------------------------------------------------------------------

linked_compartment_tracking_plot_data <-
  compartment_tracking_plot_data %>%
  filter(
    !is.na(tracking_name),
    tracking_name != ""
  ) %>%
  group_by(
    direction,
    tracking_name
  ) %>%
  filter(
    n_distinct(compartment_timepoint) >= 2
  ) %>%
  arrange(
    compartment_timepoint,
    raw_p,
    .by_group = TRUE
  ) %>%
  distinct(
    direction,
    tracking_name,
    compartment_timepoint,
    .keep_all = TRUE
  ) %>%
  ungroup()

# Label only the first appearance from left to right

linked_compartment_tracking_labels <-
  linked_compartment_tracking_plot_data %>%
  mutate(
    compartment_position =
      as.integer(compartment_timepoint)
  ) %>%
  arrange(
    direction,
    tracking_name,
    compartment_position
  ) %>%
  group_by(
    direction,
    tracking_name
  ) %>%
  slice_head(n = 1) %>%
  ungroup()


```

## Upregulated cross-compartment linkages

```{r}

# -------------------------------------------------------------------
# Upregulated matched metabolites
# -------------------------------------------------------------------

upregulated_compartment_tracking_plot <-
  linked_compartment_tracking_plot_data %>%
  filter(
    direction == "Upregulated"
  ) %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey80",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey70",
    alpha = 0.7,
    linewidth = 0.6
  ) +
  geom_point(
    aes(
      shape = supplement_label
    ),
    color = "#E15759",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_text_repel(
    data =
      linked_compartment_tracking_labels %>%
      filter(
        direction == "Upregulated"
      ),
    aes(
      label = tracking_name
    ),
    color = "#6A3D9A",
    segment.color = "#6A3D9A",
    size = 2.5,
    box.padding = 0.4,
    point.padding = 0.3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    shape = "BEP supplement"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

upregulated_compartment_tracking_plot


```

## Downregulated cross-compartment linkages

```{r}

# -------------------------------------------------------------------
# Downregulated matched metabolites
# -------------------------------------------------------------------

downregulated_compartment_tracking_plot <-
  linked_compartment_tracking_plot_data %>%
  filter(
    direction == "Downregulated"
  ) %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey80",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey70",
    alpha = 0.7,
    linewidth = 0.6
  ) +
  geom_point(
    aes(
      shape = supplement_label
    ),
    color = "#4E79A7",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_text_repel(
    data =
      linked_compartment_tracking_labels %>%
      filter(
        direction == "Downregulated"
      ),
    aes(
      label = tracking_name
    ),
    color = "#6A3D9A",
    segment.color = "#6A3D9A",
    size = 2.5,
    box.padding = 0.4,
    point.padding = 0.3,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    shape = "BEP supplement"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom"
  )

downregulated_compartment_tracking_plot


```

## Export Plots

```{r}

# -------------------------------------------------------------------
# Save plots at 16:9
# -------------------------------------------------------------------

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "upregulated_compartment_tracking_plot.png"
    ),
  plot = upregulated_compartment_tracking_plot,
  width = 16,
  height = 9,
  units = "in",
  dpi = 600,
  bg = "white"
)

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "downregulated_compartment_tracking_plot.png"
    ),
  plot = downregulated_compartment_tracking_plot,
  width = 16,
  height = 9,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

# Cross-Compartment Tracking with gglabel Repel 

## Upregulated plot

```{r}


upregulated_compartment_tracking_plot <-
  linked_compartment_tracking_plot_data %>%
  filter(
    direction == "Upregulated"
  ) %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey80",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey70",
    alpha = 0.7,
    linewidth = 0.6
  ) +
  geom_point(
    aes(
      shape = supplement_label
    ),
    color = "#E15759",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_label_repel(
    data =
      linked_compartment_tracking_labels %>%
      filter(
        direction == "Upregulated"
      ),
    aes(
      label = tracking_name
    ),
    color = "#6A3D9A",
    fill = "white",
    segment.color = "#6A3D9A",
    segment.size = 0.4,
    size = 2.5,
    label.size = 0.25,
    label.padding = unit(
      0.18,
      "lines"
    ),
    box.padding = 1.2,
    point.padding = 0.8,
    force = 15,
    force_pull = 0.05,
    min.segment.length = 0,
    max.overlaps = Inf,
    max.iter = 100000,
    max.time = 10,
    seed = 123,
    show.legend = FALSE
  ) +
  scale_y_continuous(
    expand =
      expansion(
        mult =
          c(
            0.15,
            0.35
          )
      )
  ) +
  coord_cartesian(
    clip = "off"
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    shape = "BEP supplement"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom",
    plot.margin =
      margin(
        15,
        35,
        15,
        35
      )
  )

upregulated_compartment_tracking_plot


```

## Downregulated Plot

```{r}


downregulated_compartment_tracking_plot <-
  linked_compartment_tracking_plot_data %>%
  filter(
    direction == "Downregulated"
  ) %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey80",
    linewidth = 0.4
  ) +
  geom_line(
    color = "grey70",
    alpha = 0.7,
    linewidth = 0.6
  ) +
  geom_point(
    aes(
      shape = supplement_label
    ),
    color = "#4E79A7",
    size = 2.5,
    alpha = 0.9
  ) +
  geom_label_repel(
    data =
      linked_compartment_tracking_labels %>%
      filter(
        direction == "Downregulated"
      ),
    aes(
      label = tracking_name
    ),
    color = "#6A3D9A",
    fill = "white",
    segment.color = "#6A3D9A",
    segment.size = 0.4,
    size = 2.5,
    label.size = 0.25,
    label.padding = unit(
      0.18,
      "lines"
    ),
    box.padding = 1.2,
    point.padding = 0.8,
    force = 15,
    force_pull = 0.05,
    min.segment.length = 0,
    max.overlaps = Inf,
    max.iter = 100000,
    max.time = 10,
    seed = 123,
    show.legend = FALSE
  ) +
  scale_y_continuous(
    expand =
      expansion(
        mult =
          c(
            0.35,
            0.15
          )
      )
  ) +
  coord_cartesian(
    clip = "off"
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    shape = "BEP supplement"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom",
    plot.margin =
      margin(
        15,
        35,
        15,
        35
      )
  )

downregulated_compartment_tracking_plot

```

## Export Plots

```{r}

# -------------------------------------------------------------------
# Export compartment-tracking plots at 16:9
# -------------------------------------------------------------------

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "upregulated_compartment_tracking_plot.png"
    ),
  plot = upregulated_compartment_tracking_plot,
  width = 16,
  height = 9,
  units = "in",
  dpi = 600,
  bg = "white"
)

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "downregulated_compartment_tracking_plot.png"
    ),
  plot = downregulated_compartment_tracking_plot,
  width = 16,
  height = 9,
  units = "in",
  dpi = 600,
  bg = "white"
)

```

# Distinct Colors

```{r}

# -------------------------------------------------------------------
# Create a separate color for each linked metabolite
# -------------------------------------------------------------------

tracking_names <-
  linked_compartment_tracking_plot_data %>%
  distinct(tracking_name) %>%
  arrange(tracking_name) %>%
  pull(tracking_name)

tracking_colors <-
  scales::hue_pal(
    h =
      c(
        15,
        375
      ),
    c = 100,
    l = 55
  )(length(tracking_names)) %>%
  set_names(tracking_names)

```

## Upregulated

```{r}

upregulated_compartment_tracking_plot <-
  linked_compartment_tracking_plot_data %>%
  filter(
    direction == "Upregulated"
  ) %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name,
      color = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey80",
    linewidth = 0.4
  ) +
  geom_line(
    alpha = 0.8,
    linewidth = 0.7
  ) +
  geom_point(
    aes(
      shape = supplement_label
    ),
    size = 2.5,
    alpha = 0.95
  ) +
  geom_label_repel(
    data =
      linked_compartment_tracking_labels %>%
      filter(
        direction == "Upregulated"
      ),
    aes(
      label = tracking_name
    ),
    fill = "white",
    segment.size = 0.4,
    size = 2.5,
    label.size = 0.3,
    label.padding =
      unit(
        0.18,
        "lines"
      ),
    box.padding = 1.2,
    point.padding = 0.8,
    force = 15,
    force_pull = 0.05,
    min.segment.length = 0,
    max.overlaps = Inf,
    max.iter = 100000,
    max.time = 10,
    seed = 123,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values = tracking_colors
  ) +
  scale_y_continuous(
    expand =
      expansion(
        mult =
          c(
            0.15,
            0.35
          )
      )
  ) +
  coord_cartesian(
    clip = "off"
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    shape = "BEP supplement"
  ) +
  guides(
    color = "none"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom",
    plot.margin =
      margin(
        15,
        35,
        15,
        35
      )
  )

upregulated_compartment_tracking_plot


```

## Downregulated

```{r}

downregulated_compartment_tracking_plot <-
  linked_compartment_tracking_plot_data %>%
  filter(
    direction == "Downregulated"
  ) %>%
  ggplot(
    aes(
      x = compartment_timepoint,
      y = effect_size,
      group = tracking_name,
      color = tracking_name
    )
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey80",
    linewidth = 0.4
  ) +
  geom_line(
    alpha = 0.8,
    linewidth = 0.7
  ) +
  geom_point(
    aes(
      shape = supplement_label
    ),
    size = 2.5,
    alpha = 0.95
  ) +
  geom_label_repel(
    data =
      linked_compartment_tracking_labels %>%
      filter(
        direction == "Downregulated"
      ),
    aes(
      label = tracking_name
    ),
    fill = "white",
    segment.size = 0.4,
    size = 2.5,
    label.size = 0.3,
    label.padding =
      unit(
        0.18,
        "lines"
      ),
    box.padding = 1.2,
    point.padding = 0.8,
    force = 15,
    force_pull = 0.05,
    min.segment.length = 0,
    max.overlaps = Inf,
    max.iter = 100000,
    max.time = 10,
    seed = 123,
    show.legend = FALSE
  ) +
  scale_color_manual(
    values = tracking_colors
  ) +
  scale_y_continuous(
    expand =
      expansion(
        mult =
          c(
            0.35,
            0.15
          )
      )
  ) +
  coord_cartesian(
    clip = "off"
  ) +
  labs(
    x = NULL,
    y = "Average Treatment Effect",
    shape = "BEP supplement"
  ) +
  guides(
    color = "none"
  ) +
  theme_bw(
    base_size = 8
  ) +
  theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      ),
    legend.position = "bottom",
    plot.margin =
      margin(
        15,
        35,
        15,
        35
      )
  )

downregulated_compartment_tracking_plot


```

## Export

```{r}

# -------------------------------------------------------------------
# Export plots at 16:9
# -------------------------------------------------------------------

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "upregulated_compartment_tracking_plot.png"
    ),
  plot = upregulated_compartment_tracking_plot,
  width = 16,
  height = 9,
  units = "in",
  dpi = 600,
  bg = "white"
)

ggsave(
  filename =
    file.path(
      compartment_results_root,
      "downregulated_compartment_tracking_plot.png"
    ),
  plot = downregulated_compartment_tracking_plot,
  width = 16,
  height = 9,
  units = "in",
  dpi = 600,
  bg = "white"
)

```












































































