---
title: "Exploratory Outcomes (Proteomics/UniProt)"
output: html_document
date: "2026-08-05"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(clusterProfiler) # Gene Ontology enrichment analysis
library(GO.db)           # Gene Ontology term descriptions
library(org.Hs.eg.db)    # Human gene and protein annotations
library(tidyverse)       # Data import, manipulation, and visualization
library(writexl)         # Export dataframes to Excel

```

This analysis performs Gene Ontology Biological Process over-representation analysis directly with UniProt identifiers. Each measured UniProt identifier is counted once in the foreground and background. This avoids expanding one experimentally measured protein into multiple Entrez genes.

If `GO.db` is not installed, run `BiocManager::install("GO.db")` once in the R console before knitting this document.

# Import Data

## combined_intervention_effects_results_combined_arms.RDS

```{r}

combined_intervention <-
  readRDS(
    "../1. Data/combined_intervention_effects_results_combined_arms.RDS"
  )

```

## combined_intervention_effects_results_stratified_arms.RDS

```{r}

stratified_intervention <-
  readRDS(
    "../1. Data/combined_intervention_effects_results_stratified_arms.RDS"
  )

```

# Verify Data

```{r}

proteomics_studies_timepoints <-
  bind_rows(
    combined_intervention %>%
      filter(
        description == "Proteomics"
      ) %>%
      distinct(
        study,
        studytime,
        contrast,
        measure
      ) %>%
      mutate(
        analysis_type = "Combined"
      ),

    stratified_intervention %>%
      filter(
        description == "Proteomics"
      ) %>%
      distinct(
        study,
        studytime,
        contrast,
        measure
      ) %>%
      mutate(
        analysis_type = "Stratified"
      )
  ) %>%
  select(
    analysis_type,
    study,
    studytime,
    contrast,
    measure
  ) %>%
  arrange(
    analysis_type,
    study,
    studytime,
    contrast,
    measure
  )

proteomics_studies_timepoints

```

## Compare Combined and Stratified ATEs

```{r}

proteomics_ate_comparison <-
  stratified_intervention %>%
  filter(
    description == "Proteomics",
    studytime %in%
      c(
        "Misame (1-2 mo.)",
        "Vital (2 mo.)"
      ),
    measure == "ATE"
  ) %>%
  select(
    studytime,
    biomarker,
    stratified_contrast = contrast,
    stratified_ate = est
  ) %>%
  inner_join(
    combined_intervention %>%
      filter(
        description == "Proteomics",
        studytime %in%
          c(
            "Misame (1-2 mo.)",
            "Vital (2 mo.)"
          ),
        measure == "ATE"
      ) %>%
      select(
        studytime,
        biomarker,
        combined_ate = est
      ),
    by =
      c(
        "studytime",
        "biomarker"
      )
  ) %>%
  mutate(
    ate_difference =
      stratified_ate - combined_ate,
    same_effect =
      near(
        stratified_ate,
        combined_ate
      )
  ) %>%
  arrange(
    studytime,
    stratified_contrast,
    biomarker
  )

proteomics_ate_comparison %>%
  count(
    studytime,
    stratified_contrast,
    same_effect
  )

```

The combined-arm analysis is used for MISAME-III because its single BEP/BEP intervention contrast produces the same effect estimates as the combined analysis. Contrast-specific stratified analyses are used for Mumpta-LW to distinguish BEP+ExBf from BEP+ExBf+AZT.

# Create UniProt-to-GO Mapping

## Isolate all measured proteomics UniProt identifiers

```{r}

all_proteomics_uniprot <-
  bind_rows(
    combined_intervention,
    stratified_intervention
  ) %>%
  filter(
    description == "Proteomics",
    measure == "ATE",
    !is.na(label_f),
    label_f != ""
  ) %>%
  distinct(label_f) %>%
  pull(label_f)

length(all_proteomics_uniprot)

```

## Create GO Biological Process to UniProt mapping

```{r}

uniprot_go_annotations <-
  AnnotationDbi::select(
    org.Hs.eg.db,
    keys = all_proteomics_uniprot,
    keytype = "UNIPROT",
    columns =
      c(
        "GOALL",
        "ONTOLOGYALL"
      )
  ) %>%
  as_tibble() %>%
  filter(
    ONTOLOGYALL == "BP",
    !is.na(GOALL),
    GOALL != "",
    !is.na(UNIPROT),
    UNIPROT != ""
  ) %>%
  distinct(
    GOALL,
    UNIPROT
  )

uniprot_go_term2gene <-
  uniprot_go_annotations %>%
  transmute(
    term = GOALL,
    gene = UNIPROT
  ) %>%
  distinct(
    term,
    gene
  )

uniprot_go_term2name <-
  AnnotationDbi::select(
    GO.db,
    keys =
      uniprot_go_term2gene %>%
      distinct(term) %>%
      pull(term),
    keytype = "GOID",
    columns = "TERM"
  ) %>%
  as_tibble() %>%
  transmute(
    term = GOID,
    name = TERM
  ) %>%
  filter(
    !is.na(name),
    name != ""
  ) %>%
  distinct(
    term,
    name
  )

```

## Validate UniProt mapping coverage

```{r}

uniprot_mapping_summary <-
  tibble(
    UNIPROT = all_proteomics_uniprot
  ) %>%
  mutate(
    mapped =
      UNIPROT %in% uniprot_go_term2gene$gene
  ) %>%
  count(
    mapped,
    name = "number_of_proteins"
  ) %>%
  mutate(
    percent =
      100 *
      number_of_proteins /
      sum(number_of_proteins)
  )

uniprot_mapping_summary

unmapped_proteomics_uniprot <-
  tibble(
    UNIPROT = all_proteomics_uniprot
  ) %>%
  filter(
    !UNIPROT %in% uniprot_go_term2gene$gene
  )

unmapped_proteomics_uniprot

```

# UniProt Enrichment Function

```{r}

run_uniprot_go_enrichment <-
  function(
    data,
    studytime_value,
    contrast_value,
    direction_value
  ) {
    analysis_background <-
      data %>%
      filter(
        studytime == studytime_value,
        contrast == contrast_value,
        description == "Proteomics",
        measure == "ATE",
        !is.na(label_f),
        label_f != ""
      ) %>%
      distinct(label_f) %>%
      pull(label_f) %>%
      intersect(
        uniprot_go_term2gene$gene
      )

    analysis_foreground_data <-
      data %>%
      filter(
        studytime == studytime_value,
        contrast == contrast_value,
        description == "Proteomics",
        measure == "ATE",
        sig == 1,
        !is.na(label_f),
        label_f != ""
      ) %>%
      filter(
        case_when(
          direction_value == "Downregulated" ~ est < 0,
          direction_value == "Upregulated" ~ est > 0,
          TRUE ~ FALSE
        )
      )

    analysis_foreground <-
      analysis_foreground_data %>%
      distinct(label_f) %>%
      pull(label_f) %>%
      intersect(
        analysis_background
      )

    enrichment <-
      enricher(
        gene = analysis_foreground,
        universe = analysis_background,
        pAdjustMethod = "BH",
        pvalueCutoff = 0.05,
        qvalueCutoff = 0.20,
        minGSSize = 10,
        maxGSSize = 500,
        TERM2GENE = uniprot_go_term2gene,
        TERM2NAME = uniprot_go_term2name
      )

    enrichment_results <-
      case_when(
        is.null(enrichment) ~
          list(tibble()),
        TRUE ~
          list(
            enrichment@result %>%
              as_tibble()
          )
      ) %>%
      pluck(1)

    list(
      foreground_data =
        analysis_foreground_data,
      foreground =
        analysis_foreground,
      background =
        analysis_background,
      enrichment =
        enrichment,
      results =
        enrichment_results
    )
  }

```

# Combined Analysis

## Misame

### Downregulated

#### 1-2 Months

##### Misame · Downregulated · 1-2 months · (Combined)

```{r}

misame_pn12_downregulated_uniprot <-
  run_uniprot_go_enrichment(
    data = combined_intervention,
    studytime_value = "Misame (1-2 mo.)",
    contrast_value = "BEP",
    direction_value = "Downregulated"
  )

misame3_pn12_downregulated_uniprot <-
  misame_pn12_downregulated_uniprot$foreground_data

protein_vector_misame3_pn12_downregulated_uniprot <-
  misame_pn12_downregulated_uniprot$foreground

results_misame3_pn12_downregulated_uniprot <-
  misame_pn12_downregulated_uniprot$results %>%
  mutate(
    study = "Misame",
    timepoint = "pn12",
    direction = "Downregulated",
    contrast = "BEP",
    analysis_type = "Combined",
    identifier_type = "UniProt"
  )

head(results_misame3_pn12_downregulated_uniprot)

print(results_misame3_pn12_downregulated_uniprot)

```

### Upregulated

#### 1-2 Months

##### Misame · Upregulated · 1-2 months · (Combined)

```{r}

misame_pn12_upregulated_uniprot <-
  run_uniprot_go_enrichment(
    data = combined_intervention,
    studytime_value = "Misame (1-2 mo.)",
    contrast_value = "BEP",
    direction_value = "Upregulated"
  )

misame3_pn12_upregulated_uniprot <-
  misame_pn12_upregulated_uniprot$foreground_data

protein_vector_misame3_pn12_upregulated_uniprot <-
  misame_pn12_upregulated_uniprot$foreground

results_misame3_pn12_upregulated_uniprot <-
  misame_pn12_upregulated_uniprot$results %>%
  mutate(
    study = "Misame",
    timepoint = "pn12",
    direction = "Upregulated",
    contrast = "BEP",
    analysis_type = "Combined",
    identifier_type = "UniProt"
  )

head(results_misame3_pn12_upregulated_uniprot)

print(results_misame3_pn12_upregulated_uniprot)

```

# Stratified Analysis

## Mumpta

### Downregulated

#### 2 Months

##### Mumpta · Downregulated · 2 months · BEPExBf (Stratified)

```{r}

mumpta_pn2_downregulated_bepexbf_uniprot <-
  run_uniprot_go_enrichment(
    data = stratified_intervention,
    studytime_value = "Vital (2 mo.)",
    contrast_value = "BEP+ExBf",
    direction_value = "Downregulated"
  )

mumpta_pn2_downregulated_bepexbf_stratified_uniprot <-
  mumpta_pn2_downregulated_bepexbf_uniprot$foreground_data

protein_vector_mumpta_pn2_downregulated_bepexbf_stratified_uniprot <-
  mumpta_pn2_downregulated_bepexbf_uniprot$foreground

results_mumpta_pn2_downregulated_bepexbf_stratified_uniprot <-
  mumpta_pn2_downregulated_bepexbf_uniprot$results %>%
  mutate(
    study = "Mumpta",
    timepoint = "pn2",
    direction = "Downregulated",
    contrast = "BEP+ExBf",
    analysis_type = "Stratified",
    identifier_type = "UniProt"
  )

head(results_mumpta_pn2_downregulated_bepexbf_stratified_uniprot)

print(results_mumpta_pn2_downregulated_bepexbf_stratified_uniprot)

```

##### Mumpta · Downregulated · 2 months · BEPExBfAZT (Stratified)

```{r}

mumpta_pn2_downregulated_bepexbfazt_uniprot <-
  run_uniprot_go_enrichment(
    data = stratified_intervention,
    studytime_value = "Vital (2 mo.)",
    contrast_value = "BEP+ExBf+AZT",
    direction_value = "Downregulated"
  )

mumpta_pn2_downregulated_bepexbfazt_stratified_uniprot <-
  mumpta_pn2_downregulated_bepexbfazt_uniprot$foreground_data

protein_vector_mumpta_pn2_downregulated_bepexbfazt_stratified_uniprot <-
  mumpta_pn2_downregulated_bepexbfazt_uniprot$foreground

results_mumpta_pn2_downregulated_bepexbfazt_stratified_uniprot <-
  mumpta_pn2_downregulated_bepexbfazt_uniprot$results %>%
  mutate(
    study = "Mumpta",
    timepoint = "pn2",
    direction = "Downregulated",
    contrast = "BEP+ExBf+AZT",
    analysis_type = "Stratified",
    identifier_type = "UniProt"
  )

head(results_mumpta_pn2_downregulated_bepexbfazt_stratified_uniprot)

print(results_mumpta_pn2_downregulated_bepexbfazt_stratified_uniprot)

```

### Upregulated

#### 2 Months

##### Mumpta · Upregulated · 2 months · BEPExBf (Stratified)

```{r}

mumpta_pn2_upregulated_bepexbf_uniprot <-
  run_uniprot_go_enrichment(
    data = stratified_intervention,
    studytime_value = "Vital (2 mo.)",
    contrast_value = "BEP+ExBf",
    direction_value = "Upregulated"
  )

mumpta_pn2_upregulated_bepexbf_stratified_uniprot <-
  mumpta_pn2_upregulated_bepexbf_uniprot$foreground_data

protein_vector_mumpta_pn2_upregulated_bepexbf_stratified_uniprot <-
  mumpta_pn2_upregulated_bepexbf_uniprot$foreground

results_mumpta_pn2_upregulated_bepexbf_stratified_uniprot <-
  mumpta_pn2_upregulated_bepexbf_uniprot$results %>%
  mutate(
    study = "Mumpta",
    timepoint = "pn2",
    direction = "Upregulated",
    contrast = "BEP+ExBf",
    analysis_type = "Stratified",
    identifier_type = "UniProt"
  )

head(results_mumpta_pn2_upregulated_bepexbf_stratified_uniprot)

print(results_mumpta_pn2_upregulated_bepexbf_stratified_uniprot)

```

##### Mumpta · Upregulated · 2 months · BEPExBfAZT (Stratified)

```{r}

mumpta_pn2_upregulated_bepexbfazt_uniprot <-
  run_uniprot_go_enrichment(
    data = stratified_intervention,
    studytime_value = "Vital (2 mo.)",
    contrast_value = "BEP+ExBf+AZT",
    direction_value = "Upregulated"
  )

mumpta_pn2_upregulated_bepexbfazt_stratified_uniprot <-
  mumpta_pn2_upregulated_bepexbfazt_uniprot$foreground_data

protein_vector_mumpta_pn2_upregulated_bepexbfazt_stratified_uniprot <-
  mumpta_pn2_upregulated_bepexbfazt_uniprot$foreground

results_mumpta_pn2_upregulated_bepexbfazt_stratified_uniprot <-
  mumpta_pn2_upregulated_bepexbfazt_uniprot$results %>%
  mutate(
    study = "Mumpta",
    timepoint = "pn2",
    direction = "Upregulated",
    contrast = "BEP+ExBf+AZT",
    analysis_type = "Stratified",
    identifier_type = "UniProt"
  )

head(results_mumpta_pn2_upregulated_bepexbfazt_stratified_uniprot)

print(results_mumpta_pn2_upregulated_bepexbfazt_stratified_uniprot)

```

# Combine GO Enrichment Results

```{r}

all_enrich_results_uniprot <-
  bind_rows(
    results_misame3_pn12_downregulated_uniprot %>%
      mutate(
        analysis_id =
          "misame_pn12_downregulated_combined"
      ),

    results_misame3_pn12_upregulated_uniprot %>%
      mutate(
        analysis_id =
          "misame_pn12_upregulated_combined"
      ),

    results_mumpta_pn2_downregulated_bepexbf_stratified_uniprot %>%
      mutate(
        analysis_id =
          "mumpta_pn2_downregulated_bepexbf_stratified"
      ),

    results_mumpta_pn2_upregulated_bepexbf_stratified_uniprot %>%
      mutate(
        analysis_id =
          "mumpta_pn2_upregulated_bepexbf_stratified"
      ),

    results_mumpta_pn2_downregulated_bepexbfazt_stratified_uniprot %>%
      mutate(
        analysis_id =
          "mumpta_pn2_downregulated_bepexbfazt_stratified"
      ),

    results_mumpta_pn2_upregulated_bepexbfazt_stratified_uniprot %>%
      mutate(
        analysis_id =
          "mumpta_pn2_upregulated_bepexbfazt_stratified"
      )
  ) %>%
  mutate(
    pathway_result_id =
      row_number(),

    signed_fold_enrichment =
      case_when(
        direction == "Downregulated" ~
          -abs(FoldEnrichment),
        direction == "Upregulated" ~
          abs(FoldEnrichment),
        TRUE ~ NA_real_
      ),

    uniprot_list =
      geneID %>%
      str_replace_all(
        "/",
        ", "
      ),

    uniprot_count =
      case_when(
        is.na(geneID) | geneID == "" ~ 0L,
        TRUE ~
          str_count(
            geneID,
            fixed("/")
          ) + 1L
      ),

    count_match =
      Count == uniprot_count
  ) %>%
  arrange(
    study,
    timepoint,
    contrast,
    direction,
    pvalue
  )

all_enrich_results_uniprot

```

# Validate Foreground and Background Sizes

```{r}

uniprot_analysis_size_summary <-
  bind_rows(
    tibble(
      analysis_id =
        "misame_pn12_downregulated_combined",
      foreground_proteins =
        length(misame_pn12_downregulated_uniprot$foreground),
      background_proteins =
        length(misame_pn12_downregulated_uniprot$background)
    ),

    tibble(
      analysis_id =
        "misame_pn12_upregulated_combined",
      foreground_proteins =
        length(misame_pn12_upregulated_uniprot$foreground),
      background_proteins =
        length(misame_pn12_upregulated_uniprot$background)
    ),

    tibble(
      analysis_id =
        "mumpta_pn2_downregulated_bepexbf_stratified",
      foreground_proteins =
        length(mumpta_pn2_downregulated_bepexbf_uniprot$foreground),
      background_proteins =
        length(mumpta_pn2_downregulated_bepexbf_uniprot$background)
    ),

    tibble(
      analysis_id =
        "mumpta_pn2_upregulated_bepexbf_stratified",
      foreground_proteins =
        length(mumpta_pn2_upregulated_bepexbf_uniprot$foreground),
      background_proteins =
        length(mumpta_pn2_upregulated_bepexbf_uniprot$background)
    ),

    tibble(
      analysis_id =
        "mumpta_pn2_downregulated_bepexbfazt_stratified",
      foreground_proteins =
        length(mumpta_pn2_downregulated_bepexbfazt_uniprot$foreground),
      background_proteins =
        length(mumpta_pn2_downregulated_bepexbfazt_uniprot$background)
    ),

    tibble(
      analysis_id =
        "mumpta_pn2_upregulated_bepexbfazt_stratified",
      foreground_proteins =
        length(mumpta_pn2_upregulated_bepexbfazt_uniprot$foreground),
      background_proteins =
        length(mumpta_pn2_upregulated_bepexbfazt_uniprot$background)
    )
  )

uniprot_analysis_size_summary

```

# Prepare Final Results Table

```{r}

proteomics_uniprot_results <-
  all_enrich_results_uniprot %>%
  mutate(
    study =
      recode(
        study,
        "Misame" = "MISAME-III",
        "Mumpta" = "Mumpta-LW"
      ),

    timepoint =
      recode(
        timepoint,
        "pn12" = "1-2 months",
        "pn2" = "2 months"
      )
  ) %>%
  transmute(
    `Analysis ID` = analysis_id,
    `Analysis Type` = analysis_type,
    `Identifier Type` = identifier_type,
    Study = study,
    `Time Point` = timepoint,
    Contrast = contrast,
    Direction = direction,
    `Gene Ontology ID` = ID,
    `Gene Ontology Description` =
      str_to_sentence(Description),
    `Protein Ratio` = GeneRatio,
    `Background Ratio` = BgRatio,
    `Protein Count` = Count,
    `Fold Enrichment` =
      signed_fold_enrichment,
    `P-Value` = pvalue,
    FDR = p.adjust,
    `Q-Value` = qvalue,
    `UniProt ID List` =
      uniprot_list,
    `UniProt Count` =
      uniprot_count,
    `Count Match` =
      count_match
  ) %>%
  arrange(
    Study,
    `Time Point`,
    Contrast,
    Direction,
    `P-Value`
  )

proteomics_uniprot_results

```

# Summarize Corrected Analyses

```{r}

proteomics_uniprot_enrichment_summary <-
  proteomics_uniprot_results %>%
  group_by(
    `Analysis Type`,
    Study,
    `Time Point`,
    Contrast,
    Direction
  ) %>%
  summarise(
    tested_pathways =
      n(),

    nominally_significant_pathways =
      sum(
        `P-Value` < 0.05,
        na.rm = TRUE
      ),

    fdr_significant_pathways =
      sum(
        FDR < 0.05,
        na.rm = TRUE
      ),

    minimum_p =
      min(
        `P-Value`,
        na.rm = TRUE
      ),

    minimum_fdr =
      min(
        FDR,
        na.rm = TRUE
      ),

    .groups = "drop"
  )

proteomics_uniprot_enrichment_summary

```

# Audit Manuscript Pathways

```{r}

reported_pathway_audit_uniprot <-
  proteomics_uniprot_results %>%
  filter(
    str_detect(
      `Gene Ontology Description`,
      regex(
        paste(
          c(
            "RNA-templated DNA biosynthetic",
            "telomere maintenance",
            "chromatin organization",
            "nucleosome",
            "translation",
            "negative regulation of endopeptidase",
            "negative regulation of peptidase",
            "negative regulation of proteolysis",
            "humoral immune response",
            "immunoglobulin",
            "complement activation"
          ),
          collapse = "|"
        ),
        ignore_case = TRUE
      )
    )
  ) %>%
  arrange(
    `Gene Ontology Description`,
    Study,
    Contrast,
    Direction,
    `P-Value`
  )

reported_pathway_audit_uniprot

```

# Create Protein Annotation Map

```{r}

uniprot_annotation_map <-
  AnnotationDbi::select(
    org.Hs.eg.db,
    keys = all_proteomics_uniprot,
    keytype = "UNIPROT",
    columns =
      c(
        "SYMBOL",
        "GENENAME"
      )
  ) %>%
  as_tibble() %>%
  filter(
    !is.na(UNIPROT),
    UNIPROT != ""
  ) %>%
  group_by(UNIPROT) %>%
  summarise(
    SYMBOL =
      SYMBOL[!is.na(SYMBOL)] %>%
      unique() %>%
      sort() %>%
      paste(
        collapse = "; "
      ),

    GENENAME =
      GENENAME[!is.na(GENENAME)] %>%
      unique() %>%
      sort() %>%
      paste(
        collapse = "; "
      ),

    .groups = "drop"
  )

uniprot_annotation_map

```

# Audit Manuscript Proteins

```{r}

reported_protein_audit_uniprot <-
  all_enrich_results_uniprot %>%
  transmute(
    analysis_id,
    study,
    timepoint,
    contrast,
    direction,
    analysis_type,
    ID,
    Description,
    pvalue,
    p.adjust,
    UNIPROT = geneID
  ) %>%
  separate_rows(
    UNIPROT,
    sep = "/"
  ) %>%
  filter(
    UNIPROT %in%
      c(
        "P16401",
        "P62805",
        "P05155",
        "P29622",
        "Q99727",
        "P01024",
        "P13671",
        "P08603"
      )
  ) %>%
  left_join(
    uniprot_annotation_map,
    by = "UNIPROT"
  ) %>%
  distinct() %>%
  arrange(
    UNIPROT,
    study,
    contrast,
    direction,
    pvalue
  )

reported_protein_audit_uniprot

```

# Verify Affected Protein Count

```{r}

proteomics_affected_summary <-
  combined_intervention %>%
  filter(
    description == "Proteomics",
    measure == "ATE",
    !is.na(label_f),
    label_f != ""
  ) %>%
  summarise(
    total_proteins =
      n_distinct(label_f),

    affected_proteins =
      n_distinct(
        label_f[sig == 1]
      ),

    percent_affected =
      100 *
      affected_proteins /
      total_proteins
  )

proteomics_affected_summary

```

# Export Final Workbook

```{r}

proteomics_uniprot_results_folder <-
  "../3. Results/Exploratory Outcomes (Proteomics - UniProt)"

dir.create(
  proteomics_uniprot_results_folder,
  recursive = TRUE,
  showWarnings = FALSE
)

write_xlsx(
  list(
    "All GO Results" =
      proteomics_uniprot_results,

    "FDR Significant" =
      proteomics_uniprot_results %>%
      filter(
        FDR < 0.05
      ),

    "Reported Pathway Audit" =
      reported_pathway_audit_uniprot,

    "Reported Protein Audit" =
      reported_protein_audit_uniprot,

    "Analysis Summary" =
      proteomics_uniprot_enrichment_summary,

    "Analysis Sizes" =
      uniprot_analysis_size_summary,

    "Mapping Summary" =
      uniprot_mapping_summary,

    "Unmapped UniProt" =
      unmapped_proteomics_uniprot,

    "UniProt Annotation Map" =
      uniprot_annotation_map,

    "Count Mismatches" =
      proteomics_uniprot_results %>%
      filter(
        !`Count Match`
      )
  ),
  path =
    file.path(
      proteomics_uniprot_results_folder,
      "exploratory_proteomics_uniprot_go_enrichment_results.xlsx"
    )
)

```

# Session Information

```{r}

sessionInfo()

```
