---
title: "Primary Outcomes"
output: html_document
date: "2026-07-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(broom)
library(cowplot)
library(dplyr)
library(DT)
library(ggbreak)
library(ggplot2)
library(ggrepel)
library(kableExtra)
library(knitr)
library(openxlsx)
library(readxl)
library(stringr)
library(tidyr)
library(tidyverse)
library(writexl)

```

# Import Data

## combined_intervention_effects_results_combined_arms.RDS

```{r}

combined_intervention <- 
  readRDS("../1. Data/combined_intervention_effects_results_combined_arms.RDS")

```

## combined_intervention_effects_results_stratified_arms.RDS

```{r}

stratified_intervention <- 
  readRDS("../1. Data/combined_intervention_effects_results_stratified_arms.RDS")

```

# Wrangle Data

## Generate Label for Metaboanalyst

### Combined Intervention

```{r}

# Generate label for Metaboanalyst

combined_intervention <-
  combined_intervention %>%
  mutate(
    metaboanalyst_label = label_f,
    metaboanalyst_label = str_replace_all(
      metaboanalyst_label,
      c(
        "nicotinamide mononucleotide" = "Nicotinic acid mononucleotide",
        "vitamin B3 \\( expressed as NAM\\)" = "Niacinamide",
        "nudifloramide \\(niacin catabolite\\)" = "Nudifloramide",
        "free thiamin" = "Thiamine",
        "total vitamin B1 \\(expressed as thiamin\\)" = "Thiamine",
        "vitamin B2 \\(expressed as riboflavin\\)" = "Riboflavin",
        "vitamin B12 \\(expressed as cyanocobalamin\\)" = "Cyanocobalamin",
        "vitamin B6 \\(expressed as PL\\)" = "Pyridoxine",
        "^a\\.tocopherol$|^A\\.tocopherol$" = "Alpha-tocopherol",
        "^g\\.tocopherol$|^G\\.tocopherol$" = "Gamma-tocopherol",
        "^vitamin\\.a$|^Vitamin\\.a$" = "Vitamin A"
      )
    )
  )

# Verify that everything is labeled

combined_intervention %>%
  filter(outcome_group == "primary") %>%
  distinct(metaboanalyst_label) %>%
  arrange(metaboanalyst_label) %>%
  print(n = Inf)

```

### Stratified Intervention

```{r}

# Generate label for Metaboanalyst

stratified_intervention <-
  stratified_intervention %>%
  mutate(
    metaboanalyst_label = label_f,
    metaboanalyst_label = str_replace_all(
      metaboanalyst_label,
      c(
        "nicotinamide mononucleotide" = "Nicotinic acid mononucleotide",
        "vitamin B3 \\( expressed as NAM\\)" = "Niacinamide",
        "nudifloramide \\(niacin catabolite\\)" = "Nudifloramide",
        "free thiamin" = "Thiamine",
        "total vitamin B1 \\(expressed as thiamin\\)" = "Thiamine",
        "vitamin B2 \\(expressed as riboflavin\\)" = "Riboflavin",
        "vitamin B12 \\(expressed as cyanocobalamin\\)" = "Cyanocobalamin",
        "vitamin B6 \\(expressed as PL\\)" = "Pyridoxine",
        "^a\\.tocopherol$|^A\\.tocopherol$" = "Alpha-tocopherol",
        "^g\\.tocopherol$|^G\\.tocopherol$" = "Gamma-tocopherol",
        "^vitamin\\.a$|^Vitamin\\.a$" = "Vitamin A"
      )
    )
  )

# Verify that everything is labeled

stratified_intervention %>%
  filter(outcome_group == "primary") %>%
  distinct(metaboanalyst_label) %>%
  arrange(metaboanalyst_label) %>%
  print(n = Inf)

```

## Create Reference Metabolome

### Export .xslx

```{r}

primary_reference_metabolome_xlsx <-
  combined_intervention %>%
  filter(
    outcome_group == "primary",
    !str_starts(metaboanalyst_label, "Total")) %>%
  distinct(metaboanalyst_label) %>%
  arrange(metaboanalyst_label)

primary_reference_metabolome_xlsx

# write_xlsx(
#  primary_reference_metabolome_xlsx,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Primary Outcomes Reference Metabolome.xlsx")

```

The reference metabolome was exported to Excel and uploaded to the MetaboAnalyst Metabolite ID Conversion tool using “Common Name” as the specified input type (https://www.metaboanalyst.ca/MetaboAnalyst/upload/ConvertView.xhtml). All compounds were successfully matched to MetaboAnalyst identifiers. The resulting name-mapping file was downloaded as “Primary Outcomes Reference Metabolome Name Map.csv” and saved in "Dailey-Chwalibóg, Mertens et al. 2026 Science/3. Results/Primary Outcomes (MSEA ORA)//Primary Outcomes Reference Metabolome Name Map.csv".

### Import Results from Metabolite ID Conversion and Create Final .txt File

```{r}

# Import the name map

primary_outcomes_reference_metabolome_name_map <-
  read_csv(
    "../3. Results/Primary Outcomes (MSEA ORA)/Primary Outcomes Reference Metabolome Name Map.csv")

primary_outcomes_reference_metabolome_name_map

# Export the Match column as a .txt file (one compound per line)

# write.table(
#  primary_outcomes_reference_metabolome_name_map %>%
#    filter(!is.na(Match)) %>%
#    distinct(Match) %>%
#    pull(Match),
#  file = "../3. Results/Primary Outcomes (MSEA ORA)//Primary Outcomes Reference Metabolome.txt",
#  quote = FALSE,
#  row.names = FALSE,
#  col.names = FALSE
# )

```

### Import Results from Metabolite ID Conversion and Create Final .txt File (KEGG)

```{r}

# write.table(
#  primary_outcomes_reference_metabolome_name_map %>%
#    filter(!is.na(KEGG)) %>%
#   distinct(KEGG) %>%
#   pull(KEGG),
# file = "../3. Results/Primary Outcomes (MSEA ORA)//Primary Outcomes Reference Metabolome (KEGG).txt",
# quote = FALSE,
# row.names = FALSE,
# col.names = FALSE
# )

```

# Analysis

## Elicit

### Downregulated

#### 1 Month

##### Elicit · Downregulated · 1 month (Combined)

```{r}

primary_elicit_downregulated_1mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Elicit (1 mo.)",
    outcome_group == "primary",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_elicit_downregulated_1mo_combined

```

###### ✗ No statistically significant effects.
#### 5 Months

##### Elicit · Downregulated · 5 months (Combined)

```{r}

primary_elicit_downregulated_5mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Elicit (5 mo.)",
    outcome_group == "primary",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_elicit_downregulated_5mo_combined

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_elicit_downregulated_5mo_combined,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Elicit · Downregulated · 5 months (Combined)/primary_elicit_downregulated_5mo_combined.xlsx")

```

### Upregulated

#### 1 month

##### Elicit · Upregulated · 1 month (Combined)

```{r}

primary_elicit_upregulated_1mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Elicit (1 mo.)",
    outcome_group == "primary",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_elicit_upregulated_1mo_combined

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_elicit_upregulated_1mo_combined,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Elicit · Upregulated · 1 month (Combined)/primary_elicit_upregulated_1mo_combined.xlsx")

```

Instructions to Conduct an "Enrichment Analysis (Over Representation Analysis)"

Open the .xlsx file above and copy the feature names from column Y, “metaboanalyst_label.” 

Navigate to MetaboAnalyst.ca and select “Enrichment Analysis” under “Annotated Features.” Choose “Compound names” as the Input Type, leave Feature Type set to “General,” and click “Submit.”

Verify that all, or nearly all, submitted queries have matching hits, then click “Proceed.” Retain the default pathway library: “Pathway-based SMPDB” (99 metabolite sets based on normal human metabolic pathways). Specify the reference metabolome by uploading:

"/Dailey-Chwalibóg, Mertens et al. 2026 Science/3. Results/Primary Outcomes (MSEA ORA)//Primary Outcomes Reference Metabolome.txt"

Submit the file and confirm that the blue notification states: “A total of ## were successfully added to the library.” Continue by clicking “Submit.” Review the enrichment-analysis results, then select “Download.zip” and save the archive in the appropriate folder. Finally, unzip the downloaded file and import the results.

Instructions to Conduct a Pathway Analysis:

Open the .xlsx file above and copy the feature names from column Y, “metaboanalyst_label.” 

Navigate to MetaboAnalyst.ca and select “Pathway Analysis” under “Annotated Features.” Leave the option as default "Compound List."

Choose “Compound names” as the Input Type, leave Feature Type set to “General,” and click “Submit.” 

Verify that all, or nearly all, submitted queries have matching hits, then click “Proceed.” 

Use the default parameters: Visualization method (Scatter plot (testing significant features); Enrichment method (Hypergeometric Test); Topology measure (Relative-betweeness Centrality).

Specify the reference metabolome by uploading:

"/Dailey-Chwalibóg, Mertens et al. 2026 Science/3. Results/Primary Outcomes (MSEA ORA)//Primary Outcomes Reference Metabolome (KEGG).txt"

Select the pathway library Mammals [21] Homo sapiens (SMPDB) and click proceed. Sometimes this network will throw the following error, if few metabolite features are found (ErrorThere is something wrong with the pathway enrichment analysis: Possible causes of error (last one being the most relevant): Cannot perform enrichment analysis on a single metabolite set!). In this case, select the pathway library Mammals [21] Homo sapiens (SMPDB) and see if results are generated. 

Review the enrichment-analysis results, then select “Download.zip” and save the archive in the appropriate folder. Finally, unzip the downloaded file and import the results.

Note: A minimum of 3 features is required for the Pathway Analysis module, or the following message will appear "Less than 3 metabolites left - too few to perform meaningful enrichment analysis!"

#### 5 months

##### Elicit · Upregulated · 5 months (Combined)

```{r}

primary_elicit_upregulated_5mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Elicit (5 mo.)",
    outcome_group == "primary",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_elicit_upregulated_5mo_combined

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_elicit_upregulated_5mo_combined,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Elicit · Upregulated · 5 months (Combined)/primary_elicit_upregulated_5mo_combined.xlsx")

```

## Misame

### Downregulated

#### 14-21 Days

##### Misame · Downregulated · 14-21 days · BEPBEP (Stratified)

```{r}

primary_misame_downregulated_1421d_bepbep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    contrast == "BEP/BEP",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_1421d_bepbep_stratified

```

####### ✗ No statistically significant effects.

##### Misame · Downregulated · 14-21 days · BEPIFA (Stratified)

```{r}

primary_misame_downregulated_1421d_bepifa_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    contrast == "BEP/IFA",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_1421d_bepifa_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_downregulated_1421d_bepifa_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 14-21 days · BEPIFA (Stratified)/primary_misame_downregulated_1421d_bepifa_stratified.xlsx")

```

Note: For this analysis, the only statistically significant metabolite is Arsenic. This returns the following error from MetaboAnalyst: "No match was found to the selected metabolite set library!"

##### Misame · Downregulated · 14-21 days · IFABEP (Stratified)

```{r}

primary_misame_downregulated_1421d_ifabep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    contrast == "IFA/BEP",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_1421d_ifabep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_downregulated_1421d_ifabep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 14-21 days · IFABEP (Stratified)/primary_misame_downregulated_1421d_ifabep_stratified.xlsx")

```

#### 1-2 Months

##### Misame · Downregulated · 1-2 months · BEPBEP (Stratified)

```{r}

primary_misame_downregulated_12mo_bepbep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    contrast == "BEP/BEP",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_12mo_bepbep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_downregulated_12mo_bepbep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 1-2 months · BEPBEP (Stratified)/primary_misame_downregulated_12mo_bepbep_stratified.xlsx")

```

##### Misame · Downregulated · 1-2 months · BEPIFA (Stratified)

```{r}

primary_misame_downregulated_12mo_bepifa_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    contrast == "BEP/IFA",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_12mo_bepifa_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_downregulated_12mo_bepifa_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 1-2 months · BEPIFA (Stratified)/primary_misame_downregulated_12mo_bepifa_stratified.xlsx")

```

##### Misame · Downregulated · 1-2 months · IFABEP (Stratified)

```{r}

primary_misame_downregulated_12mo_ifabep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    contrast == "IFA/BEP",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_12mo_ifabep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_downregulated_12mo_ifabep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 1-2 months · IFABEP (Stratified)/primary_misame_downregulated_12mo_ifabep_stratified.xlsx")

```

#### 3-4 Months

##### Misame · Downregulated · 3-4 months · BEPBEP (Stratified)

```{r}

primary_misame_downregulated_34mo_bepbep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    contrast == "BEP/BEP",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_34mo_bepbep_stratified

```

####### Export .xslx

```{r}

#write_xlsx(
#  primary_misame_downregulated_34mo_bepbep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 3-4 months · BEPBEP (Stratified)/primary_misame_downregulated_34mo_bepbep_stratified.xlsx")

```

##### Misame · Downregulated · 3-4 months · BEPIFA (Stratified)

```{r}

primary_misame_downregulated_34mo_bepifa_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    contrast == "BEP/IFA",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_34mo_bepifa_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_downregulated_34mo_bepifa_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Downregulated · 3-4 months · BEPIFA (Stratified)/primary_misame_downregulated_34mo_bepifa_stratified.xlsx")

```

##### Misame · Downregulated · 3-4 months · IFABEP (Stratified)

```{r}

primary_misame_downregulated_34mo_ifabep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    contrast == "IFA/BEP",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_downregulated_34mo_ifabep_stratified

```

####### ✗ No statistically significant effects.

### Upregulated

#### 14-21 Days

##### Misame · Upregulated · 14-21 days · BEPBEP (Stratified)

```{r}

primary_misame_upregulated_1421d_bepbep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    contrast == "BEP/BEP",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_1421d_bepbep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_1421d_bepbep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 14-21 days · BEPBEP (Stratified)/primary_misame_upregulated_1421d_bepbep_stratified.xlsx")

```

##### Misame · Upregulated · 14-21 days · BEPIFA (Stratified)

```{r}

primary_misame_upregulated_1421d_bepifa_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    contrast == "BEP/IFA",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_1421d_bepifa_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_1421d_bepifa_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 14-21 days · BEPIFA (Stratified)/primary_misame_upregulated_1421d_bepifa_stratified.xlsx")

```

##### Misame · Upregulated · 14-21 days · IFABEP (Stratified)

```{r}

primary_misame_upregulated_1421d_ifabep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    contrast == "IFA/BEP",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_1421d_ifabep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_1421d_ifabep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 14-21 days · IFABEP (Stratified)/primary_misame_upregulated_1421d_ifabep_stratified.xlsx")

```

#### 1-2 Months

##### Misame · Upregulated · 1-2 months · BEPBEP (Stratified)

```{r}

primary_misame_upregulated_12mo_bepbep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    contrast == "BEP/BEP",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_12mo_bepbep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
# primary_misame_upregulated_12mo_bepbep_stratified,
# path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 1-2 months · BEPBEP (Stratified)/primary_misame_upregulated_12mo_bepbep_stratified.xlsx")

```

##### Misame · Upregulated · 1-2 months · BEPIFA (Stratified)

```{r}

primary_misame_upregulated_12mo_bepifa_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    contrast == "BEP/IFA",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_12mo_bepifa_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_12mo_bepifa_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 1-2 months · BEPIFA (Stratified)/primary_misame_upregulated_12mo_bepifa_stratified.xlsx")

```

##### Misame · Upregulated · 1-2 months · IFABEP (Stratified)

```{r}

primary_misame_upregulated_12mo_ifabep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    contrast == "IFA/BEP",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_12mo_ifabep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_12mo_ifabep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 1-2 months · IFABEP (Stratified)/primary_misame_upregulated_12mo_ifabep_stratified.xlsx")

```

#### 3-4 Months

##### Misame · Upregulated · 3-4 months · BEPBEP (Stratified)

```{r}

primary_misame_upregulated_34mo_bepbep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    contrast == "BEP/BEP",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_34mo_bepbep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_34mo_bepbep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 3-4 months · BEPBEP (Stratified)/primary_misame_upregulated_34mo_bepbep_stratified.xlsx")

```

##### Misame · Upregulated · 3-4 months · BEPIFA (Stratified)

```{r}

primary_misame_upregulated_34mo_bepifa_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    contrast == "BEP/IFA",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_34mo_bepifa_stratified

```

####### ✗ No statistically significant effects.

##### Misame · Upregulated · 3-4 months · IFABEP (Stratified)

```{r}

primary_misame_upregulated_34mo_ifabep_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    contrast == "IFA/BEP",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_misame_upregulated_34mo_ifabep_stratified

```

####### Export .xslx

```{r}

# write_xlsx(
#  primary_misame_upregulated_34mo_ifabep_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Misame · Upregulated · 3-4 months · IFABEP (Stratified)/primary_misame_upregulated_34mo_ifabep_stratified.xlsx")

```

## Mumpta

### Downregulated

#### 1.5 Months

##### MumptaLW · Downregulated · 1.5 months (Combined)

```{r}

primary_mumpta_downregulated_1p5mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Vital (1.5 mo.)",
    outcome_group == "primary",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_mumpta_downregulated_1p5mo_combined

```

###### ✗ No statistically significant effects.

#### 2 Months

##### Mumpta · Downregulated · 2 months BEPExBfAZT (Stratified)

```{r}

primary_mumpta_downregulated_2mo_bepexbfazt_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Vital (2 mo.)",
    outcome_group == "primary",
    contrast == "BEP+ExBf+AZT",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_mumpta_downregulated_2mo_bepexbfazt_stratified

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_mumpta_downregulated_2mo_bepexbfazt_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Mumpta · Downregulated · 2 months BEPExBfAZT (Stratified)/primary_mumpta_downregulated_2mo_bepexbfazt_stratified.xlsx")

```

##### Mumpta · Downregulated · 2 months BEPExBf (Stratified)

```{r}

primary_mumpta_downregulated_2mo_bepexbf_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Vital (2 mo.)",
    outcome_group == "primary",
    contrast == "BEP+ExBf",
    pval < 0.05,
    est < 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_mumpta_downregulated_2mo_bepexbf_stratified

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_mumpta_downregulated_2mo_bepexbf_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Mumpta · Downregulated · 2 months BEPExBf (Stratified)/primary_mumpta_downregulated_2mo_bepexbf_stratified.xlsx")

```

### Upregulated

#### 1.5 Months

##### MumptaLW · Upregulated · 1.5 months (Combined)

```{r}

primary_mumpta_upregulated_1p5mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Vital (1.5 mo.)",
    outcome_group == "primary",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_mumpta_upregulated_1p5mo_combined

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_mumpta_upregulated_1p5mo_combined,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Mumpta · Upregulated · 1.5 months (Combined)/primary_mumpta_upregulated_1p5mo_combined.xlsx")

```

#### 2 Months

##### Mumpta · Upregulated · 2 months BEPExBfAZT (Stratified)

```{r}

primary_mumpta_upregulated_2mo_bepexbfazt_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Vital (2 mo.)",
    outcome_group == "primary",
    contrast == "BEP+ExBf+AZT",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_mumpta_upregulated_2mo_bepexbfazt_stratified

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_mumpta_upregulated_2mo_bepexbfazt_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Mumpta · Upregulated · 2 months BEPExBfAZT (Stratified)/primary_mumpta_upregulated_2mo_bepexbfazt_stratified.xlsx")

```

##### Mumpta · Upregulated · 2 months BEPExBf (Stratified)

```{r}

primary_mumpta_upregulated_2mo_bepexbf_stratified <-
  stratified_intervention %>%
  filter(
    studytime == "Vital (2 mo.)",
    outcome_group == "primary",
    contrast == "BEP+ExBf",
    pval < 0.05,
    est > 0,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )
  
primary_mumpta_upregulated_2mo_bepexbf_stratified

```

###### Export .xslx

```{r}

# write_xlsx(
#  primary_mumpta_upregulated_2mo_bepexbf_stratified,
#  path = "../3. Results/Primary Outcomes (MSEA ORA)//Mumpta · Upregulated · 2 months BEPExBf (Stratified)/primary_mumpta_upregulated_2mo_bepexbf_stratified.xlsx")

```

# Import Results

## Metabolite Site Enrichment Analyses (MSEA) · Over Representation Analyses (ORA)  

### Results From MetaboAnalyst

```{r}

# Define the main results folder

msea_root <-
  paste0(
    "../../",
    "Dailey-Chwalibóg, Mertens et al. 2026 Science/3. Results/",
    "Primary Outcomes (MSEA ORA)"
  )

# Find every msea_ora_result.csv file

msea_files <-
  msea_root %>%
  list.files(
    pattern = "^msea_ora_result[.]csv$",
    recursive = TRUE,
    full.names = TRUE
  )

# Check the number of files found

length(msea_files)

# Import files and extract metadata from folder names

primary_msea_ora_results <-
  tibble(file = msea_files) %>%
  mutate(
    
    # Analysis folder immediately above the Download folder
    source_folder =
      file %>%
      str_remove(fixed(paste0(msea_root, "/"))) %>%
      str_remove(fixed("/Download/msea_ora_result.csv")),
    
    # Study
    study =
      source_folder %>%
      str_extract("^[^·]+") %>%
      str_trim(),
    
    # Direction of regulation
    regulation =
      source_folder %>%
      str_extract("(?<=· )(Downregulated|Upregulated|Upnregulated)") %>%
      str_replace("Upnregulated", "Upregulated"),
    
    # Combined or stratified analysis
    analysis_type =
      source_folder %>%
      str_extract("(?<=\\()(Combined|Stratified)(?=\\)$)"),
    
    # Time point
    timepoint =
      case_when(
        study == "Elicit" ~
          source_folder %>%
          str_extract("[0-9.]+ months?"),
        
        study == "Misame" ~
          source_folder %>%
          str_extract("[0-9]+-[0-9]+ (days|months)"),
        
        study == "Mumpta" ~
          source_folder %>%
          str_extract("[0-9.]+ months?"),
        
        TRUE ~ NA_character_
      ),
    
    # Extract compact contrast labels
    contrast =
      case_when(
        study == "Misame" ~
          source_folder %>%
          str_extract("(BEPBEP|BEPIFA|IFABEP)"),
        
        study == "Mumpta" ~
          source_folder %>%
          str_extract("(BEPExBfAZT|BEPExBf)"),
        
        TRUE ~ NA_character_
      ),
    
    # Format contrast labels
    contrast =
      recode(
        contrast,
        "BEPBEP" = "BEP/BEP",
        "BEPIFA" = "BEP/IFA",
        "IFABEP" = "IFA/BEP",
        "BEPExBf" = "BEP+ExBf",
        "BEPExBfAZT" = "BEP+ExBf+AZT"
      ),
    
    # Import each CSV file
    data =
      map(
        file,
        ~ read_csv(
          .x,
          show_col_types = FALSE,
          name_repair = "unique_quiet"
        )
      )
  ) %>%
  
  unnest(cols = data) %>%
  
  rename(
    pathway = `...1`,
    raw_p = `Raw p`,
    holm_p = `Holm p`,
    fdr = FDR
  ) %>%
  
  select(
    study,
    timepoint,
    regulation,
    contrast,
    analysis_type,
    pathway,
    total,
    expected,
    hits,
    raw_p,
    holm_p,
    fdr,
    source_folder,
    file
  ) %>%
  
  arrange(
    study,
    timepoint,
    regulation,
    contrast,
    raw_p,
    pathway
  )

# -------------------------------------------------------------------
# Inspect the final dataframe
# -------------------------------------------------------------------

primary_msea_ora_results

glimpse(primary_msea_ora_results)

# -------------------------------------------------------------------
# Validate imported files and rows
# -------------------------------------------------------------------

primary_msea_ora_results %>%
  summarise(
    number_of_files = n_distinct(file),
    number_of_rows = n()
  )

primary_msea_ora_results %>%
  count(
    study,
    timepoint,
    regulation,
    contrast,
    analysis_type,
    name = "pathway_rows"
  ) %>%
  arrange(
    study,
    timepoint,
    regulation,
    contrast
  ) %>%
  print(n = Inf)

```

### 1. Label only FDR statistically significant

```{r}

# -------------------------------------------------------------------
# Colours
# -------------------------------------------------------------------

study_colors <-
  c(
    "ELICIT" = "#4E79A7",
    "MISAME-III" = "#F28E2B",
    "Mumpta-LW" = "#E15759"
  )

p_line_col <- "#BAB0AC"

# -------------------------------------------------------------------
# Thresholds
# -------------------------------------------------------------------

p_thr <- 0.05

y_p_line <-
  -log10(p_thr)

# -------------------------------------------------------------------
# Prepare data
# -------------------------------------------------------------------

primary_msea_ora_plot_df <-
  primary_msea_ora_results %>%
  filter(!is.na(raw_p),
         !is.na(expected),
         expected > 0,
         hits > 0) %>%
  mutate( # Calculate enrichment ratio
    enrichment_ratio =
      hits / expected,
    
    # Downregulated pathways appear on the negative side
    signed_enrichment_ratio =
      case_when(
        regulation == "Downregulated" ~ -enrichment_ratio,
        regulation == "Upregulated" ~ enrichment_ratio,
        TRUE ~ NA_real_
      ),
    
    # Calculate -log10(P-value)
    logp =
      -log10(
        pmax(
          raw_p,
          .Machine$double.xmin
        )
      ),
    
    # Format study names
    study =
      recode(
        study,
        "Elicit" = "ELICIT",
        "Misame" = "MISAME-III",
        "Mumpta" = "Mumpta-LW"
      ),
    
    study =
      factor(
        study,
        levels =
          c(
            "ELICIT",
            "MISAME-III",
            "Mumpta-LW"
          )
      ),
    
    # Abbreviate time points
    timepoint_clean =
      timepoint %>%
      str_replace_all(
        "days",
        "d"
      ) %>%
      str_replace_all(
        "months",
        "m"
      ) %>%
      str_squish(),
    
    # Add contrast where relevant
    analysis_label =
      case_when(
        is.na(contrast) ~ timepoint_clean,
        TRUE ~
          paste0(
            timepoint_clean,
            ", ",
            contrast
          )
      ),
    
    # Label FDR-significant pathways with at least two hits
    pathway_label =
      case_when(
        fdr < 0.05 & hits >= 2 ~
          paste0(
            pathway,
            " (",
            analysis_label,
            ")"
          ),
        TRUE ~ NA_character_
      ),
    
    # Grey for non-significant results
    color_group =
      case_when(
        raw_p < p_thr ~ as.character(study),
        TRUE ~ "Not significant"
      )
  )

# -------------------------------------------------------------------
# Complete colour scale
# -------------------------------------------------------------------

full_colors <-
  c(
    "Not significant" = p_line_col,
    study_colors
  )

# -------------------------------------------------------------------
# Plot
# -------------------------------------------------------------------

primary_msea_ora_plot <-
  ggplot(
    primary_msea_ora_plot_df,
    aes(
      x = signed_enrichment_ratio,
      y = logp,
      color = color_group
    )
  ) +
  
  # P-value threshold
  geom_hline(
    yintercept = y_p_line,
    color = p_line_col,
    linewidth = 0.5
  ) +
  
  # Separate downregulated and upregulated results
  geom_vline(
    xintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  
  # Pathway points
  geom_point(
    aes(
      size = hits
    ),
    alpha = 0.85
  ) +
  
  # FDR-significant pathway labels
  geom_text_repel(
    aes(
      label = pathway_label
    ),
    size = 2,
    color = "black",
    box.padding = 0.3,
    point.padding = 0.3,
    segment.color = "grey50",
    segment.size = 0.4,
    max.overlaps = Inf,
    na.rm = TRUE
  ) +
  
  scale_color_manual(
    values = full_colors
  ) +
  
  scale_size_continuous(
    range = c(
      1.5,
      4
    )
  ) +
  
  labs(
    x = "Enrichment Ratio",
    y = expression(
      -log[10](italic(P))
    ),
    color = "Study",
    size = "Hits"
  ) +
  
  theme_bw(
    base_size = 7
  ) +
  
  theme(
    axis.text = element_text(
      size = 7
    ),
    axis.title = element_text(
      size = 7
    ),
    legend.text = element_text(
      size = 7
    ),
    legend.title = element_text(
      size = 7
    ),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.margin = margin(
      5,
      5,
      5,
      5
    )
  ) +
  
  guides(
    color =
      guide_legend(
        override.aes =
          list(
            size = 3
          )
      )
  )

primary_msea_ora_plot
```

#### 

```{r}
# -------------------------------------------------------------------
# Colours
# -------------------------------------------------------------------

study_colors <-
  c(
    "ELICIT" = "#4E79A7",
    "MISAME-III" = "#F28E2B",
    "Mumpta-LW" = "#E15759"
  )

p_line_col <- "#BAB0AC"

# -------------------------------------------------------------------
# Threshold
# -------------------------------------------------------------------

p_thr <- 0.05
y_p_line <- -log10(p_thr)

# -------------------------------------------------------------------
# Prepare data
# -------------------------------------------------------------------

primary_msea_ora_plot_df <-
  primary_msea_ora_results %>%
  filter(
    !is.na(raw_p),
    !is.na(expected),
    expected > 0,
    hits > 0
  ) %>%
  mutate(
    enrichment_ratio = hits / expected,
    
    signed_enrichment_ratio =
      case_when(
        regulation == "Downregulated" ~ -enrichment_ratio,
        regulation == "Upregulated" ~ enrichment_ratio,
        TRUE ~ NA_real_
      ),
    
    logp =
      -log10(
        pmax(raw_p, .Machine$double.xmin)
      ),
    
    study =
      recode(
        study,
        "Elicit" = "ELICIT",
        "Misame" = "MISAME-III",
        "Mumpta" = "Mumpta-LW"
      ),
    
    study =
      factor(
        study,
        levels = c(
          "ELICIT",
          "MISAME-III",
          "Mumpta-LW"
        )
      ),
    
    timepoint_clean =
      timepoint %>%
      str_replace_all("days", "d") %>%
      str_replace_all("months", "m") %>%
      str_squish(),
    
    analysis_label =
      case_when(
        is.na(contrast) ~ timepoint_clean,
        TRUE ~ paste0(timepoint_clean, ", ", contrast)
      ),
    
    # Label every nominally significant result
    pathway_label =
      case_when(
        raw_p < p_thr ~
          paste0(
            pathway,
            " (",
            analysis_label,
            ")"
          ),
        TRUE ~ NA_character_
      ),
    
    color_group =
      case_when(
        raw_p < p_thr ~ as.character(study),
        TRUE ~ "Not significant"
      )
  )

# -------------------------------------------------------------------
# Identify repeated significant pathways for ellipses
# -------------------------------------------------------------------

primary_msea_ora_ellipse_df <-
  primary_msea_ora_plot_df %>%
  filter(raw_p < p_thr) %>%
  add_count(
    pathway,
    name = "pathway_count"
  ) %>%
  filter(pathway_count >= 2)

# -------------------------------------------------------------------
# Complete colour scale
# -------------------------------------------------------------------

full_colors <-
  c(
    "Not significant" = p_line_col,
    study_colors
  )

# -------------------------------------------------------------------
# Plot
# -------------------------------------------------------------------

primary_msea_ora_plot <-
  ggplot(
    primary_msea_ora_plot_df,
    aes(
      x = signed_enrichment_ratio,
      y = logp,
      color = color_group
    )
  ) +
  
  # P-value threshold
  geom_hline(
    yintercept = y_p_line,
    color = p_line_col,
    linewidth = 0.5
  ) +
  
  # Divide downregulated and upregulated results
  geom_vline(
    xintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  
  # Ellipses around repeated significant pathways
  ggforce::geom_mark_ellipse(
    data = primary_msea_ora_ellipse_df,
    aes(group = pathway),
    color = "black",
    fill = NA,
    linewidth = 0.35,
    expand = grid::unit(2, "mm"),
    inherit.aes = TRUE,
    show.legend = FALSE
  ) +
  
  # Pathway points
  geom_point(
    aes(size = hits),
    alpha = 0.85
  ) +
  
  # Labels for all nominally significant results
  geom_text_repel(
    data =
      primary_msea_ora_plot_df %>%
      filter(raw_p < p_thr),
    aes(label = pathway_label),
    size = 2,
    color = "black",
    box.padding = 0.4,
    point.padding = 0.4,
    segment.color = "grey50",
    segment.size = 0.4,
    min.segment.length = 0,
    max.overlaps = Inf,
    seed = 123,
    na.rm = TRUE
  ) +
  
  scale_color_manual(
    values = full_colors
  ) +
  
  scale_size_continuous(
    range = c(1.5, 4)
  ) +
  
  labs(
    x = "Enrichment Ratio",
    y = expression(-log[10](italic(P))),
    color = "Study",
    size = "Hits"
  ) +
  
  theme_bw(base_size = 7) +
  
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 7),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  guides(
    color =
      guide_legend(
        override.aes = list(size = 3)
      )
  )

primary_msea_ora_plot
```

# Common Pathways

```{r}

# -------------------------------------------------------------------
# Colours
# -------------------------------------------------------------------

study_colors <-
  c(
    "ELICIT" = "#4E79A7",
    "MISAME-III" = "#F28E2B",
    "Mumpta-LW" = "#E15759"
  )

p_line_col <- "#BAB0AC"

# -------------------------------------------------------------------
# Threshold
# -------------------------------------------------------------------

p_thr <- 0.05
y_p_line <- -log10(p_thr)

# -------------------------------------------------------------------
# Prepare data
# -------------------------------------------------------------------

primary_msea_ora_plot_df <-
  primary_msea_ora_results %>%
  filter(
    !is.na(raw_p),
    !is.na(expected),
    expected > 0,
    hits > 0
  ) %>%
  mutate(
    # Calculate enrichment ratio
    enrichment_ratio = hits / expected,
    
    # Downregulated pathways appear on the negative side
    signed_enrichment_ratio =
      case_when(
        regulation == "Downregulated" ~ -enrichment_ratio,
        regulation == "Upregulated" ~ enrichment_ratio,
        TRUE ~ NA_real_
      ),
    
    # Calculate -log10(P-value)
    logp =
      -log10(
        pmax(raw_p, .Machine$double.xmin)
      ),
    
    # Format study names
    study =
      recode(
        study,
        "Elicit" = "ELICIT",
        "Misame" = "MISAME-III",
        "Mumpta" = "Mumpta-LW"
      ),
    
    study =
      factor(
        study,
        levels = c(
          "ELICIT",
          "MISAME-III",
          "Mumpta-LW"
        )
      ),
    
    # Abbreviate time points
    timepoint_clean =
      timepoint %>%
      str_replace_all("days", "d") %>%
      str_replace_all("months", "m") %>%
      str_squish(),
    
    # Add contrast where relevant
    analysis_label =
      case_when(
        is.na(contrast) ~ timepoint_clean,
        TRUE ~ paste0(timepoint_clean, ", ", contrast)
      ),
    
    # Create labels for statistically significant pathways
    pathway_label =
      case_when(
        raw_p < p_thr ~
          paste0(
            pathway,
            " (",
            analysis_label,
            ")"
          ),
        TRUE ~ NA_character_
      ),
    
    # Grey for non-significant results
    color_group =
      case_when(
        raw_p < p_thr ~ as.character(study),
        TRUE ~ "Not significant"
      )
  )

# -------------------------------------------------------------------
# Identify repeated significant pathways
# -------------------------------------------------------------------

primary_msea_ora_ellipse_df <-
  primary_msea_ora_plot_df %>%
  filter(raw_p < p_thr) %>%
  add_count(
    pathway,
    name = "pathway_count"
  ) %>%
  filter(pathway_count >= 2)

# -------------------------------------------------------------------
# Complete colour scale
# -------------------------------------------------------------------

full_colors <-
  c(
    "Not significant" = p_line_col,
    study_colors
  )

# -------------------------------------------------------------------
# Plot
# -------------------------------------------------------------------

primary_msea_ora_plot <-
  ggplot(
    primary_msea_ora_plot_df,
    aes(
      x = signed_enrichment_ratio,
      y = logp,
      color = color_group
    )
  ) +
  
  # P-value threshold
  geom_hline(
    yintercept = y_p_line,
    color = p_line_col,
    linewidth = 0.5
  ) +
  
  # Separate downregulated and upregulated results
  geom_vline(
    xintercept = 0,
    color = "grey70",
    linewidth = 0.4
  ) +
  
  # Ellipses around repeated significant pathways
  ggforce::geom_mark_ellipse(
    data = primary_msea_ora_ellipse_df,
    aes(group = pathway),
    color = "black",
    fill = NA,
    linewidth = 0.35,
    expand = grid::unit(2, "mm"),
    inherit.aes = TRUE,
    show.legend = FALSE
  ) +
  
  # Pathway points
  geom_point(
    aes(size = hits),
    alpha = 0.85
  ) +
  
  # Labels only for pathways enclosed by ellipses
  geom_text_repel(
    data = primary_msea_ora_ellipse_df,
    aes(label = pathway_label),
    size = 2,
    color = "black",
    box.padding = 0.4,
    point.padding = 0.4,
    segment.color = "grey50",
    segment.size = 0.4,
    min.segment.length = 0,
    max.overlaps = Inf,
    seed = 123,
    na.rm = TRUE
  ) +
  
  scale_color_manual(
    values = full_colors
  ) +
  
  scale_size_continuous(
    range = c(1.5, 4)
  ) +
  
  labs(
    x = "Enrichment Ratio",
    y = expression(-log[10](italic(P))),
    color = "Study",
    size = "Hits"
  ) +
  
  theme_bw(base_size = 7) +
  
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 7),
    legend.position = "bottom",
    legend.box = "vertical",
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  guides(
    color =
      guide_legend(
        override.aes = list(size = 3)
      )
  )

primary_msea_ora_plot

# -------------------------------------------------------------------
# Save figure
# -------------------------------------------------------------------

ggsave(
  filename =
    "../3. Results/Primary Outcomes (MSEA ORA)/primary_msea_ora_plot.png",
  plot = primary_msea_ora_plot,
  width = 2796,
  height = 1290,
  units = "px",
  dpi = 600,
  bg = "white"
)
```

```



































































































