---
title: "Primary Outcomes (Pathway Analysis)"
output: html_document
date: "2026-08-03"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(knitr)
library(readr)
library(stringr)
library(tidyr)
library(tidyverse)
library(writexl)

```

# Import Data

## Combined intervention effects

```{r}

combined_intervention <-
  readRDS("../1. Data/combined_intervention_effects_results_combined_arms.RDS")

```

# Wrangle Data

## Generate labels for MetaboAnalyst

```{r}

combined_intervention <-
  combined_intervention %>%
  mutate(
    metaboanalyst_label = label_f,
    metaboanalyst_label = str_replace_all(
      metaboanalyst_label,
      c(
        "nicotinamide mononucleotide" = "Nicotinic acid mononucleotide",
        "vitamin B3 \\( expressed as NAM\\)" = "Niacinamide",
        "nudifloramide \\(niacin catabolite\\)" = "Nudifloramide",
        "free thiamin" = "Thiamine",
        "total vitamin B1 \\(expressed as thiamin\\)" = "Thiamine",
        "vitamin B2 \\(expressed as riboflavin\\)" = "Riboflavin",
        "vitamin B12 \\(expressed as cyanocobalamin\\)" = "Cyanocobalamin",
        "vitamin B6 \\(expressed as PL\\)" = "Pyridoxine",
        "^a\\.tocopherol$|^A\\.tocopherol$" = "Alpha-tocopherol",
        "^g\\.tocopherol$|^G\\.tocopherol$" = "Gamma-tocopherol",
        "^vitamin\\.a$|^Vitamin\\.a$" = "Vitamin A"
      )
    )
  )

combined_intervention %>%
  filter(outcome_group == "primary") %>%
  distinct(metaboanalyst_label) %>%
  arrange(metaboanalyst_label) %>%
  print(n = Inf)

```

## Create Reference Metabolome

The reference metabolome is not recreated here because it was generated previously in `Primary Outcomes (MSEA ORA).Rmd`. The same reference metabolome is used for all pathway analyses and is saved at:

`/Dailey-Chwalibóg, Mertens et al. 2026 Science/3. Results/Primary Outcomes (MSEA ORA)/Primary Outcomes Reference Metabolome.txt`

# Analysis

All analyses use the combined intervention-effect estimates and retain primary outcomes with `sigFDR == 1`. Positive and negative estimates are analyzed together.

## ELICIT

### 1 month

```{r}

primary_pathanaly_elicit_1mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Elicit (1 mo.)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_elicit_1mo_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Elicit · 1 month (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_elicit_1mo_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Elicit · 1 month (Combined)/primary_pathanaly_elicit_1mo_combined.xlsx")

```

### 5 months

```{r}

primary_pathanaly_elicit_5mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Elicit (5 mo.)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_elicit_5mo_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Elicit · 5 months (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_elicit_5mo_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Elicit · 5 months (Combined)/primary_pathanaly_elicit_5mo_combined.xlsx")

```

## MISAME-III

### 14–21 days

```{r}

primary_pathanaly_misame_1421d_combined <-
  combined_intervention %>%
  filter(
    studytime == "Misame (14-21 days)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_misame_1421d_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Misame · 14-21 days (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_misame_1421d_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Misame · 14-21 days (Combined)/primary_pathanaly_misame_1421d_combined.xlsx")

```

### 1–2 months

```{r}

primary_pathanaly_misame_12mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Misame (1-2 mo.)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_misame_12mo_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Misame · 1-2 months (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_misame_12mo_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Misame · 1-2 months (Combined)/primary_pathanaly_misame_12mo_combined.xlsx")

```

### 3–4 months

```{r}

primary_pathanaly_misame_34mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Misame (3-4 mo.)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_misame_34mo_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Misame · 3-4 months (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_misame_34mo_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Misame · 3-4 months (Combined)/primary_pathanaly_misame_34mo_combined.xlsx")

```

## Mumpta-LW

### 1.5 months

```{r}

primary_pathanaly_mumpta_1p5mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Vital (1.5 mo.)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_mumpta_1p5mo_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Mumpta · 1.5 months (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_mumpta_1p5mo_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Mumpta · 1.5 months (Combined)/primary_pathanaly_mumpta_1p5mo_combined.xlsx")

```

### 2 months

```{r}

primary_pathanaly_mumpta_2mo_combined <-
  combined_intervention %>%
  filter(
    studytime == "Vital (2 mo.)",
    outcome_group == "primary",
    sigFDR == 1,
    measure == "ATE",
    !str_starts(metaboanalyst_label, "Total")
  )

primary_pathanaly_mumpta_2mo_combined

```

#### Export `.xlsx`

```{r}

dir.create(
  "../3. Results/Primary Outcomes (Pathway Analysis)/Mumpta · 2 months (Combined)",
  recursive = TRUE,
  showWarnings = FALSE
)

# write_xlsx(
#  primary_pathanaly_mumpta_2mo_combined,
#  path = "../3. Results/Primary Outcomes (Pathway Analysis)/Mumpta · 2 months (Combined)/primary_pathanaly_mumpta_2mo_combined.xlsx")

```

# Import Results

```{r}

# Define the main results folder

pathanaly_root <-
  "../3. Results/Primary Outcomes (Pathway Analysis)"

# Find every pathway_results.csv file

pathanaly_files <-
  pathanaly_root %>%
  list.files(
    pattern = "^pathway_results[.]csv$",
    recursive = TRUE,
    full.names = TRUE
  )

length(pathanaly_files)

# Import and combine all pathway-analysis results

primary_pathanaly_results <-
  tibble(file = pathanaly_files) %>%
  mutate(
    source_folder =
      file %>%
      str_remove(fixed(paste0(pathanaly_root, "/"))) %>%
      str_remove(fixed("/Download/pathway_results.csv")),
    
    study =
      source_folder %>%
      str_extract("^[^·]+") %>%
      str_trim(),
    
    timepoint =
      source_folder %>%
      str_extract(
        "[0-9.]+(?:-[0-9.]+)? (?:days?|months?)"
      ),
    
    analysis_type =
      source_folder %>%
      str_extract(
        "(?<=\\()(Combined|Stratified)(?=\\)$)"
      ),
    
    data =
      map(
        file,
        ~ read_csv(
          .x,
          show_col_types = FALSE,
          name_repair = "unique_quiet"
        ) %>%
          rename(
            pathway = 1,
            total = Total,
            expected = Expected,
            hits = Hits,
            raw_p = `Raw p`,
            logp = `-log10(p)`,
            holm_p = `Holm adjust`,
            fdr = FDR,
            impact = Impact
          )
      )
  ) %>%
  unnest(cols = data) %>%
  select(
    study,
    timepoint,
    analysis_type,
    pathway,
    total,
    expected,
    hits,
    raw_p,
    logp,
    holm_p,
    fdr,
    impact,
    source_folder,
    file
  ) %>%
  arrange(
    study,
    timepoint,
    raw_p,
    pathway
  )

# Inspect final dataframe

primary_pathanaly_results

glimpse(primary_pathanaly_results)

# Validate imported files and rows

primary_pathanaly_results %>%
  summarise(
    number_of_files = n_distinct(file),
    number_of_rows = n()
  )

primary_pathanaly_results %>%
  count(
    study,
    timepoint,
    analysis_type,
    name = "pathway_rows"
  ) %>%
  arrange(
    study,
    timepoint
  ) %>%
  print(n = Inf)

```

# Generate Figure

```{r}

library(ggrepel)
library(ggforce)

# -------------------------------------------------------------------
# Colours and thresholds
# -------------------------------------------------------------------

study_colors <-
  c(
    "ELICIT" = "#4E79A7",
    "MISAME-III" = "#F28E2B",
    "Mumpta-LW" = "#E15759"
  )

p_line_col <- "#BAB0AC"
q_line_col <- "#59A14F"

p_thr <- 0.05
y_p_line <- -log10(p_thr)

# Highest raw P-value that remains FDR significant

fdr_p_thr <-
  primary_pathanaly_results %>%
  filter(fdr < 0.05) %>%
  summarise(fdr_p_thr = max(raw_p, na.rm = TRUE)) %>%
  pull(fdr_p_thr)

y_fdr_line <- -log10(fdr_p_thr)

# -------------------------------------------------------------------
# Prepare plot data
# -------------------------------------------------------------------

primary_pathanaly_plot_df <-
  primary_pathanaly_results %>%
  filter(
    !is.na(raw_p),
    !is.na(impact)
  ) %>%
  mutate(
    study =
      recode(
        study,
        "Elicit" = "ELICIT",
        "Misame" = "MISAME-III",
        "Mumpta" = "Mumpta-LW"
      ),
    
    study =
      factor(
        study,
        levels = c(
          "ELICIT",
          "MISAME-III",
          "Mumpta-LW"
        )
      ),
    
    timepoint_clean =
      timepoint %>%
      str_replace_all("days", "d") %>%
      str_replace_all("months", "m") %>%
      str_squish(),
    
    pathway_short =
      recode(
        pathway,
        "Nicotinate and nicotinamide metabolism" = "NAD/NAM Met",
        "Riboflavin metabolism" = "Riboflavin Met",
        "Thiamine metabolism" = "Thiamine Met",
        "Vitamin B6 metabolism" = "Vit B6 Met",
        .default = pathway
      ),
    
    pathway_label =
      case_when(
        raw_p < p_thr ~
          paste0(
            pathway_short,
            " (",
            timepoint_clean,
            ")"
          ),
        TRUE ~ NA_character_
      ),
    
    color_group =
      case_when(
        raw_p < p_thr ~ as.character(study),
        TRUE ~ "Not significant"
      )
  )

# -------------------------------------------------------------------
# Identify repeated significant pathways
# -------------------------------------------------------------------

primary_pathanaly_ellipse_df <-
  primary_pathanaly_plot_df %>%
  filter(raw_p < p_thr) %>%
  add_count(
    pathway,
    name = "pathway_count"
  ) %>%
  filter(pathway_count >= 2)

# -------------------------------------------------------------------
# Complete colour scale
# -------------------------------------------------------------------

full_colors <-
  c(
    "Not significant" = p_line_col,
    study_colors
  )

# -------------------------------------------------------------------
# Plot
# -------------------------------------------------------------------

primary_pathanaly_plot <-
  ggplot(
    primary_pathanaly_plot_df,
    aes(
      x = impact,
      y = logp,
      color = color_group
    )
  ) +
  
  geom_hline(
    yintercept = y_p_line,
    color = p_line_col,
    linewidth = 0.5
  ) +
  
  geom_hline(
    yintercept = y_fdr_line,
    color = q_line_col,
    linewidth = 0.5
  ) +
  
  annotate(
    "text",
    x = 0,
    y = y_p_line + 0.12,
    label = "italic(P)*\"-value\" < 0.05",
    parse = TRUE,
    hjust = 0,
    size = 2,
    color = p_line_col
  ) +
  
  annotate(
    "text",
    x = 0,
    y = y_fdr_line + 0.12,
    label = "italic(Q)[FDR] < 0.05",
    parse = TRUE,
    hjust = 0,
    size = 2,
    color = q_line_col
  ) +
  
  ggforce::geom_mark_ellipse(
    data = primary_pathanaly_ellipse_df,
    aes(group = pathway),
    color = "black",
    fill = NA,
    linewidth = 0.35,
    expand = grid::unit(2, "mm"),
    inherit.aes = TRUE,
    show.legend = FALSE
  ) +
  
  geom_point(
    size = 1.8,
    alpha = 0.9
  ) +
  
  geom_label_repel(
    data = primary_pathanaly_ellipse_df,
    aes(
      label = pathway_label,
      color = study
    ),
    size = 2,
    fill = "white",
    label.size = 0.2,
    box.padding = 0.35,
    point.padding = 0.3,
    segment.size = 0.4,
    min.segment.length = 0,
    max.overlaps = Inf,
    seed = 123,
    show.legend = FALSE
  ) +
  
  scale_color_manual(
    values = full_colors
  ) +
  
  scale_x_continuous(
    breaks = seq(0, 1, 0.25),
    limits = c(-0.05, 1.05)
  ) +
  
  labs(
    x = "Pathway Impact",
    y = expression(-log[10](italic(P)*"-value")),
    color = "Study"
  ) +
  
  theme_bw(base_size = 7) +
  
  theme(
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 7),
    legend.position = "bottom",
    legend.box = "horizontal",
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  guides(
    color =
      guide_legend(
        override.aes = list(size = 3)
      )
  )

primary_pathanaly_plot

# -------------------------------------------------------------------
# Save figure
# -------------------------------------------------------------------

ggsave(
  filename =
    "../3. Results/Primary Outcomes (Pathway Analysis)/primary_pathanaly_plot.png",
  plot = primary_pathanaly_plot,
  width = 240,
  height = 135,
  units = "mm",
  dpi = 600,
  bg = "white"
)

```


















































































